### 2024-07-05 Grazioli_JGRA_2024.tex
Initial Overleaf Import

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%
% Step 1: Set the \documentclass
%
%

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

%\draftfalse

%% Enter journal name below.
%% Choose from this list of Journals:
%
% JGR: Atmospheres
% JGR: Biogeosciences
% JGR: Earth Surface
% JGR: Oceans
% JGR: Planets
% JGR: Solid Earth
% JGR: Space Physics
% Global Biogeochemical Cycles
% Geophysical Research Letters
% Paleoceanography and Paleoclimatology
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resources Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%
% ie, \journalname{Water Resources Research}

% 12 Publication Units (500 words or one figure)
% 1 Table: list of distributions
% 2 Image: illustration of orientation
% 3 Image (appendix): multi-panel comparison of performance
% 4 Image distributions (filtered, sheltered/unsheltered)
% 5 image standard deviation and wind

\journalname{JGR: Atmospheres}

\begin{document}

\title{Observation of falling snow orientation in sheltered and unsheltered observation sites}

\authors{J. Grazioli\affil{1}, M. Condolf\affil{2}, F. Coletti\affil{2}, and A. Berne\affil{1}  }

\affiliation{1}{Environmental Remote Sensing Laboratory, École Polytechnique Fédérale de Lausanne, Lausanne, Switzerland}
\affiliation{2}{Department of Mechanical and Process Engineering, ETH, Zürich, Switzerland}
% \affiliation{3}{Third Affiliation}

\correspondingauthor{Jacopo Grazioli}{jacopo.grazioli@epfl.ch}

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

\begin{keypoints}
\item Observations of falling snowflake orientations show symmetrical distributions around 0°, aligning with some previous literature assumptions.
\item Sheltered and unsheltered sites show differences in orientation distributions, influenced by wind. Instrument (local perturbations) and atmospheric effects cannot be separated.
\item Numerical simulations reveal previously reported non-zero median orientations as artifacts, identifying better estimation methods.
\end{keypoints}

\begin{abstract}
This study investigates the orientation of falling snowflakes using data from a Multi-Angle Snowflake Camera (MASC). We explore the impact of different observational setups (sheltered vs. unsheltered), wind conditions, hydrometeor type, and axis ratio on the orientation distributions. Numerical simulations are used to select the best orientation estimator and to understand the reason behind different results reported in past literature. We find that previously reported non-zero median orientations are likely artifacts of averaging absolute values from individual cameras. Observed orientations generally follow a symmetrical distribution around 0$^\circ$, with broader and flatter distributions observed in unsheltered sites and high wind conditions.  Observed distributions may vary significantly from those assumed in previous studies, highlighting the need for further research on snowflake orientation under varying environmental conditions.
\end{abstract}

\section*{Plain Language Summary}
Understanding the orientation of falling snowflakes is important for interpreting weather radar data and better understand snowfall micro-physics. This study uses a 3-view camera to capture detailed images of falling snowflakes and estimate their orientations. Snowflakes generally fall with their largest dimension parallel to the ground (horizontal orientation), with variations in both directions around this setup. We found that observations in sites exposed to the wind show broader distributions of orientations compared to those in sheltered areas. Additionally, our simulations suggest that some previous studies may have incorrectly reported average orientations because of the way data was processed. These findings help refine our knowledge of snowflake behavior and show the need of better understanding the physical mechanisms affecting the distribution of snowflake orientations.

\section{Introduction}
Ice-phase hydrometeors, individual crystals and aggregates fall with a preferred orientation; In stable conditions, they align in a way that maximizes the area projected in the direction of flow~\cite{Koebschall_EF_2023}. This fact is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. Except for relatively rare situations, like storm electrification~\cite<e.g.,>{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology rely on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally.
The concept of orientation is thus linked to the anisotropic nature of most hydrometeors and it loses any meaning for isotropic, spherical particles.

Considering a population of hydrometeors rather than individual particles, as it is often the case for practical applications, the distribution of orientations is often modeled with a relatively narrow variability around 0$^\circ$ (i.e., horizontally oriented) values. Looking at the overview of Table~\ref{table:orientations}, one can observe that modeling studies mostly employ a Gaussian distribution, with standard deviations varying between 3$^\circ$~\cite<e.g.>{Matrosov_JAM_2001} and 60$^\circ$~\cite<e.g.>{Putnam_MWR_2017} depending on the type of hydrometeors.  
Observational studies~\cite<as in>{Garrett_GRL_2015} confirmed that there are differences due to hydrometeor type but also due to the environmental conditions (wind and turbulence). Observational studies report non-zero mean and median orientation values, which is contradictory with what stated so far. As we will discuss later in this manuscript, this discrepancy is in reality the result of the way multi-view instrumental observations are combined to retrieve an individual estimate of orientation.

The problem of observing hydrometeor orientations while they fall, without generating disturbance induced by the measurement itself, is not trivial. Among the ground-based instruments able in principle to perform this task we may cite: the two-dimentional video disdrometer~\cite<2DVD>{Kruger_JAOT_2002}, the multi-angle snowflake camera~\cite<MASC>{Garrett_AMT_2012}, the Precipitation Imaging Package~\cite<PIP>{Pettersen_atmosphere_2020} and the latest Video In Situ Snowfall Sensor~\cite<VISSS>{Maahn_AMT_2024}. All relatively recent, they differ for example for the number of simultaneous viewpoints (one for PIP, three for MASC, two for the others), resolution (order of magnitudes of tens of $\mu{m}$ for MASC and VISSS and hundreds of $\mu{m}$ for 2-DVD and PIP) or instrumental structure (compact for MASC and 2-DVD and open for PIP and VISSS). These sensors allow for direct investigations, based on the pictures of individual hydrometeors. As mentioned above, indirect retrievals  of the behaviour of large populations of particles within the sampling volume of remote-sensing instruments are also commonly employed, mostly by means of dual-polarization technologies~\cite<e.g.,>{Matrosov_JAS_2005},

In this manuscript we present an analysis of fall orientation of ice-phase hydrometeors, conducted on a large dataset collected by a MASC. We explore the effect of different siting of the instrument (sheltered or unsheltered), wind conditions, hydrometeor type and axis ratio as well as the estimator of orientation used. For the last investigation, We complement the analysis with simulations aimed at uncovering biases that may emerge with some estimation choices.

\section{Data and Methods}
We use a large dataset of MASC data which includes co-located environmental information, notably  temperature, humidity, wind, from the MASCdb database described in~\cite{Grazioli_SD_2022}. The database currently includes about one million of image triplets (e.g. Fig.~\ref{fig:triplet}) and pre-computed geometrical and microphysical descriptors. In this study, we filter the initial database to: (i) remove precipitation falling at positive temperatures, as we focus on dry snow; (ii) we remove particles classified as \textit{small particles} according to the classification of~\cite{Praz_AMT_2017} as their size is too close to the resolution of the images  (about 35$\mu$m); (iii) we remove all the measurements possibly due to wind-blown snow according to~\cite{Schaer_TC_2020}. All the filtering steps are conducted on data already integrated in MASCdb, thus readily reproducible. The filtered dataset includes 412'150 particles.

While operating with three 2D images of and individual 3D object, the first issue is to obtain one estimate of orientation to assign the the object/snowflake. As snowflakes are often modeled as ellipsoids or spheroids, previous studies on the MASC used the orientation of the major axis of the best-fitting ellipsoid as a proxy for orientation in individual 2D images~\cite{Garrett_GRL_2015,Gergely_JGRA_2016, Fitch_AMT_2021, Fitch_JGR_2022}, and then taking the mean of the absolute value of the three estimates. We refer to this approach to \textit{Mean\_Abs}. Other studies perform instead a 3D reconstruction of an ellipsoid that best fits the three views, and take as estimate the orientation of the major axis of such ellipsoid~\cite{Jiang_JAS_2019}. We refer to this family of approaches as \textit{Ellipsoid\_Fit}. At a first glance in Table~\ref{table:orientations}, the results of the experimental studies using those estimators seem to disprove the common assumption that the distribution of orientations of snowflakes is centered at 0$^\circ$, and modal values between 10$^\circ$ and 20$^\circ$ are reported. While for the \textit{Ellipsoid\_Fit} in ~\cite{Jiang_JAS_2019} this is based on a small sample of aggregates, thus difficult to evaluate, we would like to show that the appearance of non-zero modal values in the \textit{Mean\_Abs} approach is a bias induced by conducting this type of averaging.

To do so, and to select an appropriate estimator of orientation, we conducted numerical simulations (results shown in Fig.~\ref{fig:simulations}). Ellipsoid of randomly varying geometry as well as simulated aggregates and monomers are generated~\cite<as for example in>{Leinonen_AMT_2021} and their projection into MASC-like images is computed in order to test various estimation methods. The orientation of the reference snowflakes or ellipsoid is individually known and it is drawn by a zero-mean and 25$^\circ$ standard deviation from a Gaussian distribution. While this simulation setup has the potential to be exploited much further than this with the variations of all the parameters involved, the goal is here to support the selection of the estimator that we will employ in the actual field-collected data. Figure~\ref{fig:simulations} allows us to have an overview of how well various methods reproduce the distribution of orientation (panel a,c) and the magnitude of the errors with respect to the absolute value of the orientation (panel b, d). We tested the following estimator: \textit{Min}, being the minimum (absolute value) orientation among the views with sign preserved, \textit{Mean\_Abs} as defined above, \textit{Ellipsoid\_fit} as described above, \textit{Mean} being the mean of the estimate of three views, as well as the individual estimates of each view \textit{Cam0}, \textit{Cam1}, \textit{Cam2}.

The results of this experiment shows first of all that the non-zero modal values of \textit{Mean\_Abs} appear also in this case, thus confirming the hypothesis that this is an artefact of averaging absolute-values, which makes near-zero estimates much less likely to occur. \textit{Ellipsoid\_fit} has small errors, but the modal values of the distribution of this method consistently show a positive bias of a few degrees, such that the reconstructed distribution is not centered exactly at 0$^\circ$. Individual camera views have larger errors as well as very broad reconstructed distributions, a result in itself interesting and worth investigating further as many snowflake imagers have only one view. \textit{Min} provides low errors, but it produces distributions that are exaggeratedly leptokurtic. Therefore in the following, we will use a simple \textit{Mean}, which is an optimal compromise between preserving as much as possible the shape of the distribution with limited errors and extreme simplicity. For this specific experimental setup, \textit{Mean} is also the only method that likely preserves Gaussianity, according to a Kolmogorov-Smirnov test.

\section{Experimental results and discussion}
The distribution of all the (\textit{Mean}) orientation values presented in Fig.~\ref{fig:alldata} shows that the overall distribution is symmetrical with a mode at 0$^\circ$. The distribution of all observations is fat-tailed. Different type of instrument installations have a clear impact on the distribution: installations within DFIRs, with enhanced wind sheltering and protection from local turbulence shows less pronounced tails while not sheltered sites more pronounced ones as well as a much flatter distribution in the proximity of 0$^\circ$ orientation values, between -25$^\circ$ and 25$^\circ$. It can be hypothesized that this change in shape of the distribution is the result of the different siting, with the unsheltered observations perturbed by \textit{local} turbulence driven by the instrument itself~\cite<as described by>{Fitch_AMT_2021}.

It is however reasonable to assume that the \textit{atmospheric} conditions should play a role in shaping the distribution of orientation values, with broader distribution in case of strong winds and wind variations. In Fig.~\ref{fig:wind-effect} we compare the evolution of the standard deviation of orientation distributions for DFIR and unsheltered conditions at various levels of horizontal winds. At low winds ($<1$ms$^{-1}$), the curves are relatively close, as well as for high winds ($>10$ms$^{-1}$) while they depart up to 5/6$^\circ$ for intermediate conditions, with the DFIR curve increasing as the wind increases. It can be assumed that for near-zero winds local turbulence due to the instrument itself is minimal, and sheltered and unsheltered conditions are substantially equivalent. It is however not straightforward to interpret the behaviour of the curves in case of stronger winds. One explanation could be that as the wind gets stronger, the DFIR shelter gets less effective, leading to a broadening of the distribution (broadening due to local interactions with the instrument). However, strong wind conditions could broaden the distribution as well (broadening due to atmospheric causes) and a combination of the two mechanisms is plausible. Further research would be needed to clarify this aspect and for the time we can simply report this discrepancy among the different siting types.

Under the assumption that measurements collected within a DFIR are inherently less perturbed, we focus now the attention only on this type of siting. Given the large size of the original dataset, we can perform further data stratification. Figure~\ref{fig:hydroclass} shows the distribution of orientations in two contrasting wind regimes ($<4$ms$^{-1}$ and $>8$ms$^{-1}$), stratified according to hydrometeor type~\cite<method of>{Praz_AMT_2017} and average axis ratio of the particles (divided in two classes of below- and above-median axis ratio).  The effect of stronger winds results in broader and platykurtic distributions, which lose the clear median around 0$^\circ$ orientation values. For columnar and planar crystals of elongated shape (low axis ratio), this results in distributions that start to exhibit off-zero modal values, that are not observed in low-wind conditions. 
The effect of axis ratio on the distribution is not remarkable, but this is an important information in itself: we can exclude that the modal values around 0$^\circ$ may be artefacts resulting from near-spherical axis-symmetrical particles being assigned a quasi-horizontal orientations merely because of geometrical conventions.

If we focus only on low-wind conditions ($<4$ms$^{-1}$) within DFIR sites, a situation that should minimize perturbations for our dataset, the distribution of orientation values has $\sigma$ of 28$^\circ$ (graupel), 30$^\circ$ (aggregates), 32$^\circ$ (planar crystals), and 34$^\circ$ (columnar crystals), 29.5$^\circ$ (all mixed). Those values are within the range of commonly assumed ones, as reported in Table~\ref{table:orientations}, but depart significantly from individual studies especially the ones assuming extremely narrow distributions~\cite{Matrosov_JAM_2001,Matrosov_JAS_2005}, or the ones assuming much broader distributions~\cite<e.g.>{Ryzhkov_JAMC_2011, Bukovic_JAMC_2018}. Further research on the actual distribution of orientations is needed to fine-tune those estimates, especially in order to understand the effect of actual atmospheric conditions, not instrument-induced, on distribution shape and breadth.

\section{Conclusions}
The main results of this letter devoted to the estimation of the orientation of falling snowflakes with a MASC are:
\begin{itemize}
    \item Estimation of orientation as the mean of the orientation of the best fit ellipsoid of each camera views is a good compromise in terms of error minimization and correct reproduction of the distribution of orientation values. 
    \item Previous studies reporting median orientations departing from 0$^\circ$ are most likely resulting from artefacts due to averaging the \textit{absolute value} of the orientation of each camera view.
    \item Single-camera estimates are largely more uncertain than a combination of the three views.
    \item Field measurements on a large dataset show that orientation values have overall a symmetrical distribution around 0$^\circ$.
    \item Observations collected in unsheltered sites as well as in sheltered sites in strong wind conditions result in broader and flatter distributions, with a less clear mode at 0$^\circ$. This is probably both an artefact of local disturbances (instrument-induced) as well as actual environmental phenomenon. It is not possible in this set-up to separate the two effects and further research is needed in this field.
    \item Observed orientation distributions can differ largely with respect to assumptions made in previous studies. 
\end{itemize}

% ---------------
 \begin{table}
 \caption{Non-exhaustive list of research works that provide assumptions or measurements of the distribution of orientation angles of ice phase particles. $\overline{\theta}$ indicates mean orientation, $\sigma$ is the standard deviation,  $\theta_{M}$ indicates the mode of the distribution, $\theta_{Q50}$ is the median, $MAD(\theta)$ is the median absolute deviation.}
 \label{table:orientations}
 \centering
 \begin{tabular}{l p{80mm}} 
 \hline
  Ref.   & Details \\
 \hline
   \cite{Matrosov_JAM_2001} & $\overline{\theta} = 0^\circ, \sigma = 3^\circ$ \\
   \hline
   \cite{Matrosov_JAS_2005} & $\overline{\theta} = 0^\circ, \sigma = 9^\circ \pm3^\circ$ (dendrites) \\
   \hline
    \cite{Kennedy_JAMC_2011}  &   $\overline{\theta} = 0^\circ, \sigma = 15^\circ$ (dendrites) \newline $\overline{\theta} = 0^\circ, \sigma = 30^\circ$ (aggregates)     \\
   \hline
    \cite{Ryzhkov_JAMC_2011} & $\overline{\theta} = 0^\circ, \sigma = 40^\circ$ (aggregates) \newline 
   $\overline{\theta} = 0^\circ \sigma = 40^\circ$ (graupel) \\
   \hline
   \cite{Hogan_JAMC_2012} & Assumed horizontal  \\
   \hline   
   \cite{Putnam_MWR_2017} & $\overline{\theta} = 0^\circ \sigma = 20^\circ$ (snow) \newline 
   $\overline{\theta} = 0^\circ \sigma \in [0,60]^\circ$ (graupel) \\
   \hline
   \cite{Bukovic_JAMC_2018} &  $\overline{\theta} = 0^\circ, \sigma = 10^\circ$ (dendrites) \newline  $\overline{\theta} = 0^\circ \sigma = 40^\circ$ (aggregates) \\
   \hline
    \cite{Matsui_JGRA_2019}  & $\overline{\theta} = 0^\circ \sigma = 20^\circ$ (aggregates) \newline 
   $\overline{\theta} = 20^\circ \sigma = 42^\circ$ (graupel) \\
   \hline
   \cite{Garrett_GRL_2015}$^*$ & $\theta_{Q50} = 39^\circ, \theta_{M} = 13^\circ$ (aggregates) \newline
   $\theta_{Q50} = 35^\circ, \theta_{M} = 16^\circ$ (rimed particles) \newline
   $\theta_{Q50} = 35^\circ, \theta_{M} = 20^\circ$ (graupel) \newline
   $\theta_{Q50} = 28^\circ, \theta_{M} = 13^\circ$ (low-turbulence, all types) \newline
   $\theta_{Q50}= 40^\circ, \theta_{M} = 23^\circ$ (high-turbulence, all types) \\
   \hline
   \cite{Gergely_JGRA_2016}$^*$ &  $\theta_{Q50} \in [38,47.3]^\circ$, $MAD(\theta) \in [43,51]^\circ$ \\
   \hline
   \cite{Jiang_JAS_2019}$^*$ & $\theta_{M}\approx 10^\circ$ \\
   \hline
   \cite{Fitch_AMT_2021}$^*$ & $\theta_{M} = 57^\circ$ (wind $> 5$ ms$^{-1}$) \newline
   $\theta_{M} = 28^\circ$ (wind $< 1.5$ ms$^{-1}$) \\
   \hline
   \cite{Fitch_JGR_2022}$^*$ & $\theta_{M} = 3^\circ$ 
   \newline
   $\theta_{M} = 13^\circ$ (different estimations)  \\
   \hline
   \cite{Schrom_JAS_2023}$^*$ & Beta distribution (varying parameters) \\
   \hline
    \multicolumn{2}{l}{$^{*}$Papers that present orientation statistics from ground-based measurements.}
 \end{tabular}
 \end{table}

% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
\begin{figure}
 \noindent\includegraphics[width=\textwidth]{Fig01.png}
\caption{Example of a triplet of images of a snow aggregate, collected by a MASC. The best fitted ellipse of each image is drawn in green, with the major axis defining the apparent orientation with respect to the horizontal.    }
\label{fig:triplet}
\end{figure}

\begin{figure}
 \noindent \centering \includegraphics[width=\textwidth]{Fig02.png}
\caption{Evaluation of the retrieval of orientation that could be achieved with a triplet of MASC images, conducted using simulated data (ellipsoids in the left column, simulated snowflakes in the right column). (a) and (c): distribution of orientation value, both reference and reconstructed. (b) and (d): letter-value plots of the distribution of the error of the estimation of the absolute value of the orientation. The simulations are conducted on 10'000 ellipsoids of various geometry as well as on 10'000 aggregates. }
\label{fig:simulations}
\end{figure}

\begin{figure}
 \noindent \centering \includegraphics[width=\textwidth]{Fig03.png}
\caption{Histogram of the distribution of orientation value for all data (412'150 particles, top) and data stratified according to the instrumental setup (bottom) either being unsheltered (274'839 particles) or deployed within a double-fence inter-comparison reference (DFIR, 137'311 particles). The dotted lines illustrate the shape of a Gaussian curve having the same mean and standard deviation as the observed data, and are given to provide a visual comparison of the difference with respect to a Gaussian shape (it is not a fit of the histogram). Orientation values are estimated as the sign-preserving mean of the three MASC views.   }
\label{fig:alldata}
\end{figure}

\begin{figure}
 \noindent \centering \includegraphics[width=\textwidth]{Fig04.png}
\caption{Evolution of the standard deviation of the  distribution of orientation values at various wind speed levels and stratified according to the instrumental setup. The bottom panel shows the number of observations falling into each bin of wind speed.  }
\label{fig:wind-effect}
\end{figure}

\begin{figure}
 \noindent \centering \includegraphics[width=\textwidth]{Fig05.png}
\caption{Violin plot (distribution) of orientation values stratified according to different hydrometeor types. The data correspond to observations collected within a DFIR (Fig.~\ref{fig:wind-effect}), for low horizontal wind (about 47'000 observations for wind $<4$ ms$^{-1}$) and high horizontal wind (about 36'000 observations for wind $>8$ ms$^{-1}$). Data are stratified in two classes of axis ratio (a\_r) limited by the median axis ratio value of the entire population. The number of observations in each distribution ranges between a minimum of 500 (planar crystals, low axis ratio, low wind) and a maximum of about 15'000 (aggregates, low axis ratio, low wind).} 
\label{fig:hydroclass}
\end{figure}

%% SIDEWAYS FIGURE and TABLE
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{figsamp}
% \caption{caption here}
% \label{newfig}
% \end{sidewaysfigure}
%
%  \begin{sidewaystable}
%  \caption{Caption here}
% \label{tab:signif_gap_clos}
%  \begin{tabular}{ccc}
% one&two&three\\
% four&five&six
%  \end{tabular}
%  \end{sidewaystable}

%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
%\begin{linenomath*}
%\begin{equation}
%y|{f} \sim g(m, \sigma),
%\end{equation}
%\end{linenomath*}

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%
% Acronyms
%   \begin{acronyms}
%   \acro{Acronym}
%   Definition here
%   \acro{EMOS}
%   Ensemble model output statistics
%   \acro{ECMWF}
%   Centre for Medium-Range Weather Forecasts
%   \end{acronyms}

\section{Open Research}
The MASCDB dataset used to reproduce the field-based results of this study is openly available in~\citeA{MASCDB_dataset_2023}

%AGU requires an Availability Statement for the underlying data needed to understand, evaluate, and build upon the reported research at the time of peer review and publication.

%Authors should include an Availability Statement for the software that has a significant impact on the research. Details and templates are in the Availability Statement section of the Data and Software for Authors Guidance: \url{https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#availability}

%It is important to cite individual datasets in this section and, and they must be included in your bibliography. Please use the type field in your bibtex file to specify the type of data cited. Some options include Dataset, Software, Collection, ComputationalNotebook. Ex:

\acknowledgments

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\bibliography{references,references_data}
%
% don't specify bibliographystyle

% In the References section, cite the data/software described in the Availability Statement (this includes primary and processed data used for your research). For details on data/software citation as well as examples, see the Data & Software Citation section of the Data & Software for Authors guidance
% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#citation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\bibliography{enter your bibtex bibliography filename here}

%Reference citation instructions and examples:
%
% Please use ONLY \cite and \citeA for reference citations.
% \cite for parenthetical references
% ...as shown in recent studies (Simpson et al., 2019)
% \citeA for in-text citations
% ...Simpson et al. (2019) have shown...
%
%
%...as shown by \citeA{jskilby}.
%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
%...has been shown \cite{jskilbye}.
%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown by \cite <e.g.,>[and others]{lewin76}.
%
% apacite uses < > for prenotes and [ ] for postnotes
% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \citealp, etc.).
% \nocite is okay to use to add references from your Supporting Information
%

\end{document}

More Information and Advice:

%% ------------------------------------------------------------------------ %%

### 2024-07-09 Grazioli_JGRA_2024.tex
Updates from Overleaf

@@
 
-\authors{J. Grazioli\affil{1}, M. Condolf\affil{2}, F. Coletti\affil{2}, and A. Berne\affil{1}  }
+%\authors{J. Grazioli\affil{1}, M. Condolf\affil{2}, F. Coletti\affil{2}, and A. Berne\affil{1}  }
 
-\affiliation{1}{Environmental Remote Sensing Laboratory, École Polytechnique Fédérale de Lausanne, Lausanne, Switzerland}
-\affiliation{2}{Department of Mechanical and Process Engineering, ETH, Zürich, Switzerland}
+%\affiliation{1}{Environmental Remote Sensing Laboratory, École Polytechnique Fédérale de Lausanne, Lausanne, Switzerland}
+%\affiliation{2}{Department of Mechanical and Process Engineering, ETH, Zürich, Switzerland}
 % \affiliation{3}{Third Affiliation}
@@
 \begin{keypoints}
-\item Observations of falling snowflake orientations show symmetrical distributions around 0°, aligning with some previous literature assumptions.
+\item Observations of snowflake orientations show symmetrical distributions around 0$^\circ$ and standard deviations that confirm some of the assumptions made in non-observational studies. 
 \item Sheltered and unsheltered sites show differences in orientation distributions, influenced by wind. Instrument (local perturbations) and atmospheric effects cannot be separated.
-\item Numerical simulations reveal previously reported non-zero median orientations as artifacts, identifying better estimation methods.
+\item Numerical simulations reveal that common estimation choices made in past studies may lead to data processing artifacts.
 \end{keypoints}
@@
 Numerical simulations are used to select the best orientation estimator and to understand the reason behind different results reported in past literature.
-We find that previously reported non-zero median orientations are likely artifacts of averaging absolute values from individual cameras.
-Observed orientations generally follow a symmetrical distribution around 0$^\circ$, with broader and flatter distributions observed in unsheltered sites and high wind conditions.
+We find that previously reported non-zero median orientations are likely artifacts of averaging absolute values of apparent orientations from individual cameras.
+Observed orientations generally follow a symmetrical distribution around 0$^\circ$, with broader and flatter distributions observed in unsheltered sites and/or high wind conditions.
 Observed distributions may vary significantly from those assumed in previous studies, highlighting the need for further research on snowflake orientation under varying environmental conditions.
@@
 \section{Introduction}
-Ice-phase hydrometeors, individual crystals and aggregates fall with a preferred orientation; In stable conditions, they align in a way that maximizes the area projected in the direction of flow~\cite{Koebschall_EF_2023}.
+
+Ice-phase hydrometeors, such as individual crystals and aggregates, fall with a preferred orientation.
+In stable conditions, they align in a way that maximizes the area projected in the direction of flow \cite{Koebschall_EF_2023}.
 This fact is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where.
-Except for relatively rare situations, like storm electrification~\cite<e.g.,>{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology rely on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally.
-The concept of orientation is thus linked to the anisotropic nature of most hydrometeors and it loses any meaning for isotropic, spherical particles.
+Except for relatively rare situations, like storm electrification \cite{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology relies on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally.
+The concept of orientation is thus linked to the anisotropic nature of most hydrometeors and loses any meaning for isotropic, spherical particles.
 
-Considering a population of hydrometeors rather than individual particles, as it is often the case for practical applications, the distribution of orientations is often modeled with a relatively narrow variability around 0$^\circ$ (i.e., horizontally oriented) values.
-Looking at the overview of Table~\ref{table:orientations}, one can observe that modeling studies mostly employ a Gaussian distribution, with standard deviations varying between 3$^\circ$~\cite<e.g.>{Matrosov_JAM_2001} and 60$^\circ$~\cite<e.g.>{Putnam_MWR_2017} depending on the type of hydrometeors.
-Observational studies~\cite<as in>{Garrett_GRL_2015} confirmed that there are differences due to hydrometeor type but also due to the environmental conditions (wind and turbulence).
-Observational studies report non-zero mean and median orientation values, which is contradictory with what stated so far.
-As we will discuss later in this manuscript, this discrepancy is in reality the result of the way multi-view instrumental observations are combined to retrieve an individual estimate of orientation.
+Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ (i.e., horizontally oriented) values.
+Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly employ a Gaussian distribution model, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors.
+Observational studies \cite{Garrett_GRL_2015} confirm that differences exist due to hydrometeor type and environmental conditions (wind and turbulence).
+Observational studies report as well non-zero mean and median orientation values, which is contradictory with what was stated previously.
+As we will discuss later in this manuscript, this discrepancy is, in reality, the result of the way multi-view instrumental observations are combined to retrieve an individual estimate of orientation.
 
 The problem of observing hydrometeor orientations while they fall, without generating disturbance induced by the measurement itself, is not trivial.
-Among the ground-based instruments able in principle to perform this task we may cite: the two-dimentional video disdrometer~\cite<2DVD>{Kruger_JAOT_2002}, the multi-angle snowflake camera~\cite<MASC>{Garrett_AMT_2012}, the Precipitation Imaging Package~\cite<PIP>{Pettersen_atmosphere_2020} and the latest Video In Situ Snowfall Sensor~\cite<VISSS>{Maahn_AMT_2024}.
-All relatively recent, they differ for example for the number of simultaneous viewpoints (one for PIP, three for MASC, two for the others), resolution (order of magnitudes of tens of $\mu{m}$ for MASC and VISSS and hundreds of $\mu{m}$ for 2-DVD and PIP) or instrumental structure (compact for MASC and 2-DVD and open for PIP and VISSS).
-These sensors allow for direct investigations, based on the pictures of individual hydrometeors.
-As mentioned above, indirect retrievals  of the behaviour of large populations of particles within the sampling volume of remote-sensing instruments are also commonly employed, mostly by means of dual-polarization technologies~\cite<e.g.,>{Matrosov_JAS_2005},
+Among the ground-based instruments designed to perform this task, we may cite the two-dimensional video disdrometer (2DVD) \cite{Kruger_JAOT_2002}, the multi-angle snowflake camera (MASC) \cite{Garrett_AMT_2012}, the Precipitation Imaging Package (PIP) \cite{Pettersen_atmosphere_2020}, and the Video In Situ Snowfall Sensor (VISSS) \cite{Maahn_AMT_2024}.
+These relatively recent instruments differ, for example, in the number of simultaneous viewpoints (one for PIP, three for MASC, two for the others), resolution (on the order of tens of micrometers for MASC and VISSS and hundreds of micrometers for 2DVD and PIP), or instrumental structure (compact for MASC and 2DVD and open-field for PIP and VISSS).
+These sensors allow for direct microphysical investigations based on pictures of individual hydrometeors.
+As mentioned above, indirect retrievals of the behavior of large populations of particles within the sampling volume of remote-sensing instruments are also commonly employed, mostly by means of dual-polarization technologies \cite{Matrosov_JAS_2005}.
 
-In this manuscript we present an analysis of fall orientation of ice-phase hydrometeors, conducted on a large dataset collected by a MASC.
-We explore the effect of different siting of the instrument (sheltered or unsheltered), wind conditions, hydrometeor type and axis ratio as well as the estimator of orientation used.
-For the last investigation, We complement the analysis with simulations aimed at uncovering biases that may emerge with some estimation choices.
-
-\section{Data and Methods}
-We use a large dataset of MASC data which includes co-located environmental information, notably  temperature, humidity, wind, from the MASCdb database described in~\cite{Grazioli_SD_2022}.
-The database currently includes about one million of image triplets (e.g.
-Fig.~\ref{fig:triplet}) and pre-computed geometrical and microphysical descriptors.
-In this study, we filter the initial database to: (i) remove precipitation falling at positive temperatures, as we focus on dry snow; (ii) we remove particles classified as \textit{small particles} according to the classification of~\cite{Praz_AMT_2017} as their size is too close to the resolution of the images  (about 35$\mu$m); (iii) we remove all the measurements possibly due to wind-blown snow according to~\cite{Schaer_TC_2020}.
-All the filtering steps are conducted on data already integrated in MASCdb, thus readily reproducible.
-The filtered dataset includes 412'150 particles.
-
-While operating with three 2D images of and individual 3D object, the first issue is to obtain one estimate of orientation to assign the the object/snowflake.
-As snowflakes are often modeled as ellipsoids or spheroids, previous studies on the MASC used the orientation of the major axis of the best-fitting ellipsoid as a proxy for orientation in individual 2D images~\cite{Garrett_GRL_2015,Gergely_JGRA_2016, Fitch_AMT_2021, Fitch_JGR_2022}, and then taking the mean of the absolute value of the three estimates.
-We refer to this approach to \textit{Mean\_Abs}.
-Other studies perform instead a 3D reconstruction of an ellipsoid that best fits the three views, and take as estimate the orientation of the major axis of such ellipsoid~\cite{Jiang_JAS_2019}.
-We refer to this family of approaches as \textit{Ellipsoid\_Fit}.
-At a first glance in Table~\ref{table:orientations}, the results of the experimental studies using those estimators seem to disprove the common assumption that the distribution of orientations of snowflakes is centered at 0$^\circ$, and modal values between 10$^\circ$ and 20$^\circ$ are reported.
-While for the \textit{Ellipsoid\_Fit} in ~\cite{Jiang_JAS_2019} this is based on a small sample of aggregates, thus difficult to evaluate, we would like to show that the appearance of non-zero modal values in the \textit{Mean\_Abs} approach is a bias induced by conducting this type of averaging.
-
-To do so, and to select an appropriate estimator of orientation, we conducted numerical simulations (results shown in Fig.~\ref{fig:simulations}).
-Ellipsoid of randomly varying geometry as well as simulated aggregates and monomers are generated~\cite<as for example in>{Leinonen_AMT_2021} and their projection into MASC-like images is computed in order to test various estimation methods.
-The orientation of the reference snowflakes or ellipsoid is individually known and it is drawn by a zero-mean and 25$^\circ$ standard deviation from a Gaussian distribution.
-While this simulation setup has the potential to be exploited much further than this with the variations of all the parameters involved, the goal is here to support the selection of the estimator that we will employ in the actual field-collected data.
-Figure~\ref{fig:simulations} allows us to have an overview of how well various methods reproduce the distribution of orientation (panel a,c) and the magnitude of the errors with respect to the absolute value of the orientation (panel b, d).
-We tested the following estimator: \textit{Min}, being the minimum (absolute value) orientation among the views with sign preserved, \textit{Mean\_Abs} as defined above, \textit{Ellipsoid\_fit} as described above, \textit{Mean} being the mean of the estimate of three views, as well as the individual estimates of each view \textit{Cam0}, \textit{Cam1}, \textit{Cam2}.
-
-The results of this experiment shows first of all that the non-zero modal values of \textit{Mean\_Abs} appear also in this case, thus confirming the hypothesis that this is an artefact of averaging absolute-values, which makes near-zero estimates much less likely to occur.
-\textit{Ellipsoid\_fit} has small errors, but the modal values of the distribution of this method consistently show a positive bias of a few degrees, such that the reconstructed distribution is not centered exactly at 0$^\circ$.
-Individual camera views have larger errors as well as very broad reconstructed distributions, a result in itself interesting and worth investigating further as many snowflake imagers have only one view.
-\textit{Min} provides low errors, but it produces distributions that are exaggeratedly leptokurtic.
-Therefore in the following, we will use a simple \textit{Mean}, which is an optimal compromise between preserving as much as possible the shape of the distribution with limited errors and extreme simplicity.
-For this specific experimental setup, \textit{Mean} is also the only method that likely preserves Gaussianity, according to a Kolmogorov-Smirnov test.
-
-\section{Experimental results and discussion}
-The distribution of all the (\textit{Mean}) orientation values presented in Fig.~\ref{fig:alldata} shows that the overall distribution is symmetrical with a mode at 0$^\circ$.
-The distribution of all observations is fat-tailed.
-Different type of instrument installations have a clear impact on the distribution: installations within DFIRs, with enhanced wind sheltering and protection from local turbulence shows less pronounced tails while not sheltered sites more pronounced ones as well as a much flatter distribution in the proximity of 0$^\circ$ orientation values, between -25$^\circ$ and 25$^\circ$.
-It can be hypothesized that this change in shape of the distribution is the result of the different siting, with the unsheltered observations perturbed by \textit{local} turbulence driven by the instrument itself~\cite<as described by>{Fitch_AMT_2021}.
-
-It is however reasonable to assume that the \textit{atmospheric} conditions should play a role in shaping the distribution of orientation values, with broader distribution in case of strong winds and wind variations.
-In Fig.~\ref{fig:wind-effect} we compare the evolution of the standard deviation of orientation distributions for DFIR and unsheltered conditions at various levels of horizontal winds.
-At low winds ($<1$ms$^{-1}$), the curves are relatively close, as well as for high winds ($>10$ms$^{-1}$) while they depart up to 5/6$^\circ$ for intermediate conditions, with the DFIR curve increasing as the wind increases.
-It can be assumed that for near-zero winds local turbulence due to the instrument itself is minimal, and sheltered and unsheltered conditions are substantially equivalent.
-It is however not straightforward to interpret the behaviour of the curves in case of stronger winds.
-One explanation could be that as the wind gets stronger, the DFIR shelter gets less effective, leading to a broadening of the distribution (broadening due to local interactions with the instrument).
-However, strong wind conditions could broaden the distribution as well (broadening due to atmospheric causes) and a combination of the two mechanisms is plausible.
-Further research would be needed to clarify this aspect and for the time we can simply report this discrepancy among the different siting types.
-
-Under the assumption that measurements collected within a DFIR are inherently less perturbed, we focus now the attention only on this type of siting.
-Given the large size of the original dataset, we can perform further data stratification.
-Figure~\ref{fig:hydroclass} shows the distribution of orientations in two contrasting wind regimes ($<4$ms$^{-1}$ and $>8$ms$^{-1}$), stratified according to hydrometeor type~\cite<method of>{Praz_AMT_2017} and average axis ratio of the particles (divided in two classes of below- and above-median axis ratio).
-The effect of stronger winds results in broader and platykurtic distributions, which lose the clear median around 0$^\circ$ orientation values.
-For columnar and planar crystals of elongated shape (low axis ratio), this results in distributions that start to exhibit off-zero modal values, that are not observed in low-wind conditions.
-The effect of axis ratio on the distribution is not remarkable, but this is an important information in itself: we can exclude that the modal values around 0$^\circ$ may be artefacts resulting from near-spherical axis-symmetrical particles being assigned a quasi-horizontal orientations merely because of geometrical conventions.
-
-If we focus only on low-wind conditions ($<4$ms$^{-1}$) within DFIR sites, a situation that should minimize perturbations for our dataset, the distribution of orientation values has $\sigma$ of 28$^\circ$ (graupel), 30$^\circ$ (aggregates), 32$^\circ$ (planar crystals), and 34$^\circ$ (columnar crystals), 29.5$^\circ$ (all mixed).
-Those values are within the range of commonly assumed ones, as reported in Table~\ref{table:orientations}, but depart significantly from individual studies especially the ones assuming extremely narrow distributions~\cite{Matrosov_JAM_2001,Matrosov_JAS_2005}, or the ones assuming much broader distributions~\cite<e.g.>{Ryzhkov_JAMC_2011, Bukovic_JAMC_2018}.
-Further research on the actual distribution of orientations is needed to fine-tune those estimates, especially in order to understand the effect of actual atmospheric conditions, not instrument-induced, on distribution shape and breadth.
-
-\section{Conclusions}
-The main results of this letter devoted to the estimation of the orientation of falling snowflakes with a MASC are:
-\begin{itemize}
-    \item Estimation of orientation as the mean of the orientation of the best fit ellipsoid of each camera views is a good compromise in terms of error minimization and correct reproduction of the distribution of orientation values.
-    \item Previous studies reporting median orientations departing from 0$^\circ$ are most likely resulting from artefacts due to averaging the \textit{absolute value} of the orientation of each camera view.
-    \item Single-camera estimates are largely more uncertain than a combination of the three views.
-    \item Field measurements on a large dataset show that orientation values have overall a symmetrical distribution around 0$^\circ$.
-    \item Observations collected in unsheltered sites as well as in sheltered sites in strong wind conditions result in broader and flatter distributions, with a less clear mode at 0$^\circ$.
-This is probably both an artefact of local disturbances (instrument-induced) as well as actual environmental phenomenon.
-It is not possible in this set-up to separate the two effects and further research is needed in this field.
-    \item Observed orientation distributions can differ largely with respect to assumptions made in previous studies. 
-\end{itemize}
+In this manuscript, we present an analysis of the fall orientation of ice-phase hydrometeors, conducted on a large dataset collected by a MASC.
+We explore the effect of different siting of the instrument (sheltered or unsheltered), wind conditions, hydrometeor type, and axis ratio, as well as the estimator of orientation used.
+For the last investigation, we complement the analysis with simulations aimed at uncovering biases that may emerge with some estimation choices.
 
@@
 
-% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
+\section{Data, Methods, Preliminary Considerations}
+
+We utilize a large dataset of MASC observations, which includes co-located environmental information such as temperature, humidity, and wind, sourced from the MASCdb database extensively described in \cite{Grazioli_SD_2022}.
+The MASC instrument~\cite{Garrett_AMT_2012} captures simultaneous images of falling snowflakes with three cameras looking at a common focal point from three different viewpoints (0$^\circ$, +36$^\circ$, -36$^\circ$).
+
+MASCdb includes approximately one million image triplets (see Fig.~\ref{fig:triplet}) and pre-computed geometrical and microphysical descriptors.
+For the purpose of the present study, we filter the initial dataset by: (i) excluding precipitation occurring at positive temperatures, focusing thus on dry snow; (ii) removing particles classified as \textit{small particles} following the criteria in \cite{Praz_AMT_2017}, due to their size being too close to the image resolution (which is about 35$\mu$m); and (iii) excluding measurements potentially affected or dominated by wind-blown snow according to \cite{Schaer_TC_2020}.
+All filtering steps are conducted on data already integrated into MASCdb, ensuring reproducible results.
+The resulting filtered dataset comprises 412'150 particles.
+
 \begin{figure}
@@
 \end{figure}
+
+In handling three 2D images of an individual 3D object, a primary challenge is to derive a single  estimate of orientation for each snowflake.
+Snowflakes are commonly modeled as ellipsoids or spheroids in previous MASC studies: the orientation of the major axis of the best-fitting ellipsoid in each 2D image is estimated, and the average of the three views is considered as a proxy for orientation \cite{Garrett_GRL_2015, Gergely_JGRA_2016, Fitch_AMT_2021, Fitch_JGR_2022}.
+The cited studies calculate the mean of the absolute value of the three estimates, an approach referred to here as \textit{Mean\_Abs}.
+Alternatively, other studies perform a 3D reconstruction of the ellipsoid that best fits all three views and take the orientation of its major axis as the estimate \cite{Jiang_JAS_2019}, a family of approaches named \textit{Ellipsoid\_Fit}.
+Examining Table~\ref{table:orientations}, the experimental studies using these estimators report the occurrence of non-zero modal orientation values between 10$^\circ$ and 20$^\circ$, challenging the common assumption of a distribution centered at 0$^\circ$.
+While the non-zero modal values reported for \textit{Ellipsoid\_Fit} in \cite{Jiang_JAS_2019} are based on a small sample of aggregate snowflakes, making evaluation difficult, we demonstrate that the appearance of non-zero modal values in the \textit{Mean\_Abs} approach results simply from averaging absolute values.
+This will offer a reconciliation between the common assumptions made on snowflake orientations and the contradiction of the first studies that used the MASC, a contradiction that was observed also by~\citeA{Garrett_GRL_2015} and discussed in the Conclusions of their works.
+
+To address this, and to select an appropriate orientation estimator, we conducted numerical simulations (results shown in Fig.~\ref{fig:simulations}).
+These simulations generated ellipsoids with randomly varying geometries, as well as more realistic and geometrically complex aggregates following methods such as those described in \cite{Leinonen_AMT_2021}.
+The objects, of known 3D structure are projected into MASC-like 2D images to test various estimation methods.
+The orientation of the reference snowflake or ellipsoids is known and drawn from a Gaussian distribution with a zero mean and 25$^\circ$ standard deviation.
+Although this simulation setup could be expanded extensively to investigate the contribution of each ingredient at play, the goal here is to support the selection of an estimator to use with actual field-collected data.
+Figure~\ref{fig:simulations} provides an overview of how various methods reproduce the orientation distribution (panels a and c) and the magnitude of errors relative to the absolute orientation value (panels b and d).
+Tested estimators include: \textit{Min}, representing the minimum (absolute value) orientation among the views with preserved sign; \textit{Mean\_Abs}, as previously defined; \textit{Ellipsoid\_Fit}, as described; \textit{Mean}, the mean of the estimates from the three views; and individual estimates from each view (\textit{Cam0}, \textit{Cam1}, \textit{Cam2}).
 
@@
 
+The result of the simulations reveal that the non-zero modal values observed with \textit{Mean\_Abs} are artifact resulting from averaging absolute values,  reducing the likelihood of near-zero estimates.
+With an example: let us assume that the individual estimate of orientation from the three cams were -10$^\circ$, 0$^\circ$ and +20$^\circ$.
+With \textit{Mean\_Abs} the final estimate would be approximately 10$^\circ$, while for \textit{Mean} it would be roughly 3$^\circ$.
+The results shown here show that, at least statistically-speaking, a bias is introduced.
+
+The \textit{Ellipsoid\_Fit} method exhibits small errors but its distribution consistently shows a positive bias of a few degrees, making it not centered precisely at 0$^\circ$.
+Individual camera views yield larger errors and broader distributions, which is in itself a noteworthy result given that many snowflake imagers only capture one view.
+\textit{Min} provides low errors but produces distributions that are excessively leptokurtic.
+Consequently, for subsequent analyses, we adopt a straightforward \textit{Mean}, which is possibly an optimal balance between preserving distribution shape with minimal errors and simplicity.
+Importantly, \textit{Mean} is the only method in this specific experimental setup that likely maintains Gaussianity,  according to a Kolmogorov-Smirnov test.
+
+\section{Experimental Results and Discussion}
+The distribution of \textit{Mean} orientation values in field-collected data, as depicted in Fig.~\ref{fig:alldata}, reveals an overall symmetrical distribution centered at 0$^\circ$.
+However, the distribution has fat tails with the frequency of large orientation values exceeding the ones of a Gaussian of equivalent standard deviation.
+The installation conditions of the instruments significantly influence this distribution: sites within Double-Fence Intercomparison Reference~\cite<DFIR, see for example>{Smith_HESS_2020}, providing enhanced wind sheltering and protection from local turbulence, show less pronounced tails.
+In contrast, unsheltered sites exhibit more pronounced tails and a flatter distribution near 0$^\circ$, ranging between -25$^\circ$ and 25$^\circ$.
+This suggests that unsheltered observations are likely perturbed by local turbulence induced by the instrument itself, as described in~\citeA{Fitch_AMT_2021}.
+
 \begin{figure}
- \noindent \centering \includegraphics[width=\textwidth]{Fig03.png}
+ \noindent \centering \includegraphics[width=0.8\textwidth]{Fig03.png}
 \caption{Histogram of the distribution of orientation value for all data (412'150 particles, top) and data stratified according to the instrumental setup (bottom) either being unsheltered (274'839 particles) or deployed within a double-fence inter-comparison reference (DFIR, 137'311 particles).
@@
 \end{figure}
+
+Atmospheric conditions are expected to play a role in shaping orientation distributions, particularly increasing wind conditions.
+Figure~\ref{fig:wind-effect} compares the standard deviation of orientation distributions between DFIR and unsheltered sites at different wind speeds.
+At low wind speeds ($<1$ ms$^{-1}$) and high wind speeds ($>10$ ms$^{-1}$), the curves are close.
+However, they diverge by up to 5/6$^\circ$ for intermediate wind speeds, with the DFIR curve showing an increase with wind speed once the wind speed exceeds 4 ms$^{-1}$ and being relatively constant below that threshold.
+Under near-zero wind conditions, local turbulence induced by the instrument geometry is minimal, making DFIR and unsheltered conditions comparable.
+One hypothesis could be that until 4 ms$^{-1}$ the DFIR is effective in minimizing local turbulence (DFIR curve constant, unsheltered curve abruptly increasing), while it gradually loses this capacity until "saturating" around 10 ms$^{-1}$ (DFIR curve increasing, unsheltered curve constant).
+Making this assumption is however implying that all the observed variations is due to the instrument itself and not a physical phenomenon of interaction of turbulent air and snowflakes.
+Further research is therefore necessary to clarify these dynamics.
 
@@
 
+Given that the measurements within DFIRs are less perturbed, we focus now on this siting type.
+With the original dataset's large size, we conduct further data stratification.
+Figure~\ref{fig:hydroclass} illustrates orientation distributions under two contrasting wind regimes ($<4$ ms$^{-1}$ and $>8$ ms$^{-1}$), stratified by hydrometeor type~\cite{Praz_AMT_2017} and average axis ratio.
+Stronger winds lead to broader and platykurtic distributions, where a clear median around 0$^\circ$ orientation values is less evident and the distribution starts to assume uniform shape.
+For elongated columnar and planar crystals (low axis ratio), this results in distributions showing off-zero modal values, which were not observed under low-wind conditions.
+
+The influence of axis ratio on distribution is not significant, but this is an important information.
+It allows to exclude that the observed distribution could be an artefact given by the convention of assigning 0$^\circ$ orientation values to particles with axis ratio of 1.
+As the shape of the distribution is preserved for low and high axis ratios, we can be confident of the physical nature of this distribution shape.
+ Focusing on low-wind conditions ($<4$ ms$^{-1}$) within DFIR sites, i.e.
+the subset with minimal external perturbations, the orientation distribution exhibits standard deviations of 28$^\circ$ (graupel), 30$^\circ$ (aggregates), 32$^\circ$ (planar crystals), and 34$^\circ$ (columnar crystals), with a combined value of 29.5$^\circ$ for all types.
+These values align with commonly assumed ranges (Table~\ref{table:orientations}), yet differ notably from individual studies employing extremely narrow or broad distributions~\cite{Matrosov_JAM_2001, Matrosov_JAS_2005, Ryzhkov_JAMC_2011, Bukovic_JAMC_2018}.
+Future research should refine these estimates, particularly to discern the impact of actual atmospheric conditions versus instrument-induced effects on distribution shape and breadth.
+
 \begin{figure}
@@
 \end{figure}
+
+\section{Conclusions}
+The main results of this study on the estimation of the orientation of falling snowflakes using a MASC are summarized as follows.
+At first, estimating orientation as the mean of the orientation of the best-fit ellipsoid from each camera view represents a reasonable approach in terms of minimizing errors and accurately reproducing the distribution of orientation values.
+Single-camera estimates exhibit high uncertainty compared to combining estimates from all three views and this should be kept in mind if single-view imagers are used.
+Previous studies reporting median orientations deviating from 0$^\circ$ in MASC studies are likely artifacts resulting from averaging the \textit{absolute value} of orientation estimates from each camera view.
+This explication, demonstrated with numerical simulations in the manuscript, is simpler than what proposed for example by~\cite{Garrett_GRL_2015} who hypothesized fluttering and local, instrument-induced, turbulence (which has anyway an impact) as possible explanations.
+Field measurements from a large dataset demonstrate that orientation values generally exhibit a symmetrical distribution around 0$^\circ$.
+Observations from unsheltered sites or in very high wind conditions result in broader and flatter orientation distributions, with a less pronounced mode at 0$^\circ$.
+This is likely influenced by both local disturbances (instrument-induced) and environmental factors, although with the current information we can only observe this discrepancy.
+It is clear that further research, with different setups is needed to separate the contributions of these effects and estimate the \textit{natural} effect of wind and turbulence on the distribution of orientation values.
+Finally, we observed that orientation distributions of some hydrometeor types can vary significantly with respect to the assumptions made in previous studies.
+
+\section{Open Research}
+The MASCDB dataset used to for the field-based data of this study is openly available in~\citeA{MASCDB_dataset_2023}.
+The scripts to reproduce the results in this manuscript are openly available at \hl{TODO: add github repo with the scripts}.
+
+% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
 
@@
 %   \end{acronyms}
-
-\section{Open Research}
-The MASCDB dataset used to reproduce the field-based results of this study is openly available in~\citeA{MASCDB_dataset_2023}
 

### 2024-11-29 Review_R1_answers_202412.tex
Updates from Overleaf

% LaTeX rebuttal letter example. 
% 
% Copyright 2019 Friedemann Zenke, fzenke.net
%
% Based on examples by Dirk Eddelbuettel, Fran and others from 
% https://tex.stackexchange.com/questions/2317/latex-style-or-macro-for-detailed-response-to-referee-report
% 
% Licensed under cc by-sa 3.0 with attribution required.
% See https://creativecommons.org/licenses/by-sa/3.0/
% and https://stackoverflow.blog/2009/06/25/attribution-required/

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage{xcolor}
\usepackage{float}
\usepackage{soul}
\usepackage{graphicx}
\usepackage{breakcites}
\usepackage{url}

% import Eq and Section references from the main manuscript where needed
% \usepackage{xr}
% \externaldocument{manuscript}

% package needed for optional arguments
\usepackage{xifthen}
\newcommand*{\blue}{\textcolor{blue}}

\begin{document}

\section*{Response to the reviewers of 2024GL113042 }

% General intro text goes here
We thank the reviewers for their critical assessment of our work, their positive feedback and very useful suggestions. 
In the following we address their concerns point by point.\\

\noindent

\begin{table}[H]
\centering
\begin{tabular}{ |c|c| } 
 \hline
 {\itshape italic text  } & {\itshape Comments of the reviewers}  \\ 
 roman text               & Our answers or comments   \\ 
 \blue{blue text}         & \blue{Verbatim quotes from the manuscript}  \\ 
 \hline
\end{tabular}
\end{table}

\noindent
A manuscript is included with highlighted differences and modifications applied during revision is submitted as well.

\subsection*{Editor Comment}
\textit{Thank you for submitting "Observation of falling snowflakes orientation at sheltered and unsheltered sites" [Paper \#2024GL113042] to Geophysical Research Letters. I have received 2 reviews of your manuscript, which are included below and/or attached. As you can see, the reviews indicate that major revisions are needed before we can consider proceeding with your paper. I am therefore returning the paper to you so that you can make the necessary changes.
}

\newpage
\subsection*{Reviewer 1}

\textit{The manuscript is well-written. The authors did a great job addressing the observational issues with the Multi-Angel Snowflake Camera (MASC). I particularly appreciate the reverse-engineering numerical experiments: idealized and realistic objects were generated to evaluate various methods for describing the orientation of those objects. The findings from the numerical experiment, together with the analysis of the in-situ data, will be a valuable addition to the literature and a useful resource for the remote sensing community. I recommend the manuscript be published in GRL, after the following issues are addressed:}

\subsubsection*{Major comments:}
\begin{enumerate}
    \item \textit{The discussion about the impact of the environmental winds on the orientation of the snowflakes is not sound. Wind speed alone may not be sufficient to capture this impact. The distribution of wind components likely plays a more significant role.}
\end{enumerate}

\subsubsection*{Minor comments:}
\begin{enumerate}
    \item \textit{Including results on graupel in a study focused specifically on snowflakes raises questions about the sampling method.}
\end{enumerate}

\clearpage
\subsection*{Reviewer 2}
\textit{The authors present an interesting analysis of the orientation behavior of ice crystals and snowflakes during falling. The analysis is based on a large database of ground-based in-situ observations (Multi-Angle Snowflake Camera). Dependencies of the orientation on observational setups (wind mitigation), environmental parameters and particle properties are discussed. A simulation experiment further reveals important impacts of using different methods to estimate the average orientation angle from multiple cameras.}\\

\textit{Overall, I find the paper very well written and also the non-expert reader can easily follow. I have only a number of minor comments and suggestions. One might of course question whether the topic is of interest to a wider community, as it is a rather specific topic. However, I see the topic itself as highly relevant for active and passive microwave remote sensing but also for estimation of ice particle sedimentation velocity. Therefore, I recommend the manuscript for publication in GRL.}

\subsubsection*{Comments:}

\begin{enumerate}

    \item \textit{There is a very recent laboratory study about ice particle orientation which should be added to reference list and I also highly suggest to discuss their results. They appear to be slightly contradictory to your study:} \cite{Stout_ACP_2024} \\

    We thank the reviewer for this, extremely relevant, recommendation.  This high quality study, with laboratory experiments using snowflake replicas in a liquid solution shows that planar crystals may be falling with regimes that do not have a preferential orientation around 0$^\circ$. We believe that, given some of the assumptions of~\cite{Stout_ACP_2024}, their results are somehow complementary to our work and point out to the need to better understand the role of turbulence in orientation of snow particles. Regarding the cited study, we have the following comments:
    \begin{itemize}
        \item A relevant limitation of this work is that "it  only considers quiescent conditions, as the authors want to know under what conditions particle instability still occurs, even without the addition of turbulence.". This is quite different with respect to our field-based experiments.
        \item The study focuses on planar crystals and it uses 12 sub-types of this  hydrometeors (from plates to dendrites). In our case, the planar crystals are all classified together into an individual class, which is also one of the most difficult to classify with the MASC and to observe (given the 35 microns resolution of the MASC itself). In this sense our study is coarser and cannot go to this level of detail, which could be a reason why we do not observe such behavior. 
        \item Interestingly, another recent study~\cite{Koebschall_EF_2023} using replicas of snow aggregates and a liquid solution, showed instead that the particles tend to orient such that the area projected in the direction of flow is maximized, which is the type of behavior that is consistent with our observations. This is very interesting as two separate indoor studies, with similar methods, seem to indicate that individual crystals of planar type and aggregate snowflakes may be behaving differently regarding preferential orientation.
        \item It is unfortunate that the study reports, as possible backup evidence and as driving motivation, that previous studies using MASC also showed off-zero preferential orientations. This is exactly what we showed, in our paper, to be due to the bias of averaging \textbf{absolute value} of orientation estimates taken from different viewpoints and it was one of the driving motivation of our publication.  
    \end{itemize}

    We now modified the manuscript to give some space to commenting and elaborating on the findings of~\cite{Stout_ACP_2024}. At first, we include this study in Table~1.

    \item \textit{L. 37: Not sure if "standard deviation" is a term that is suited for the plain language summary. Maybe one can formulate simpler: "typical variability of the orientation angle\dots"}\\

    \noindent
    We agree with the reviewer, and we implemented their suggestion. \\

    \item Introduction: Honestly, comparing it with the other sections, I find the introduction a bit weak. Considering the wider audience addressed by GRL, I think the importance of ice particle orientation should be emphasized and explained more. It is certainly true that most radar polarimetric quantities are sensitive to orientation (L. 50-52). But this is by far not all in my opinion. Particle orientation also impacts the scattering signal in passive microwave (MW) observations (both from space and from ground). Many space-borne MW sensors are also polarization-sensitive. Considering the enormous efforts undertaken for example by ECMWF and other services to assimilate those all-sky MW observations, this aspect should be mentioned and cited. Another aspect is that an ice particle falling with a preferred orientation will have a very different sedimentation velocity. So characterizing the particle orientation is quite key to simulate its terminal velocity (which is quite an important parameter in any cloud model). Ice particle orientation is also relevant for determining cloud radiative properties such as optical depth or albedo.

    \item \textit{L. 46: It would be good to add a reference for this statement. I question a bit whether this general statement is actually true. I completely agree for plate-like crystals or large aggregates. But what is for example about bullet rosettes in high cirrus clouds? Or columns/needles? Do we have solid knowledge about their falling orientation? Also the references you provide a few sentences later (\cite{Noel_JAMC_2005,Tinklenberg_JFM_2023}) focus only on planar crystals.}\\

    \item \textit{L. 47: What do you mean by "relatively simple"? Personally, I find all those ice particles terribly complicated.}\\

    In this case our intention was to differentiate between individual crystals (the shape of which can indeed be complicated, but that usually have a clear symmetry) from aggregates and graupel, which are composed by many crystals with additional contribution of riming, making it hard to understand the individual shapes or define a clear geometry. Our phrasing was confusing and this sentence needed to be reformulated in any case (see previous points). It now reads:

     \hl{TO DO}

    \item \textit{Table 1: I like the idea of having a table summarizing previous studies. But are there no studies that looked at columns or needles?}\\

    \cite{Klett_JAS_1995, Reinking_JAMC_1997}

    \item \textit{L. 140: Parenthesis around the reference Leinonen et al., 2021 should be removed.}

    \noindent
    It is now corrected.\\

    \item \textit{L. 157: Why "approximately"?}

    \noindent
    The reviewer is right. The sentence now reads:\\
    \blue{With \textit{Mean\_Abs} the final estimate would be 10$^\circ$, while for \textit{Mean} it would be 3.33$^\circ$}\\

    \item L. 211-215: One general thought regarding the fact that radar studies most often "retrieve" smaller range of orientation angles: If one looks at vertically pointing radar images, one always finds the lowest few hundred meters much more turbulent than the ice and snow cloud aloft. Surface friction is certainly responsible for this effect. So I am wondering if the often observable low turbulence in many ice clouds (for example indicated by small spectrum width) could be the reason why the orientation angle distribution is more narrow? Of course, the discrepancy might also be connected to unrealistic scattering properties often assumed by polarimetric shape retrievals.

    \item Connected to this topic but even more complicated to answer is why some polarimetric radar observations seem to indicate a certain non-zero distribution of the particle orientation: \cite{Melnikov_JAOT_2013}
\end{enumerate}

\bibliographystyle{apalike}
\bibliography{./references.bib}

\end{document}

### 2024-11-29 Grazioli_JGRA_2024_R1.tex
Updates from Overleaf

--- 2024-07-09 Grazioli_JGRA_2024.tex
@@
 
-\title{Observation of falling snow orientation in sheltered and unsheltered observation sites}
+\title{Observation of falling snowflakes orientation at sheltered and unsheltered sites}
 
-%\authors{J. Grazioli\affil{1}, M. Condolf\affil{2}, F. Coletti\affil{2}, and A. Berne\affil{1}  }
+\authors{J. Grazioli\affil{1}, M. Condolf\affil{2}, Y.-A. Roulet\affil{3} F. Coletti\affil{2}, and A. Berne\affil{1}  }
 
-%\affiliation{1}{Environmental Remote Sensing Laboratory, École Polytechnique Fédérale de Lausanne, Lausanne, Switzerland}
-%\affiliation{2}{Department of Mechanical and Process Engineering, ETH, Zürich, Switzerland}
-% \affiliation{3}{Third Affiliation}
+\affiliation{1}{Environmental Remote Sensing Laboratory, École Polytechnique Fédérale de Lausanne, Lausanne, Switzerland}
+\affiliation{2}{Department of Mechanical and Process Engineering, ETH, Zürich, Switzerland}
+\affiliation{3}{Swiss Federal Office of Meteorology and Climatology (MeteoSwiss)}
 
@@
 \begin{keypoints}
-\item Observations of snowflake orientations show symmetrical distributions around 0$^\circ$ and standard deviations that confirm some of the assumptions made in non-observational studies. 
-\item Sheltered and unsheltered sites show differences in orientation distributions, influenced by wind. Instrument (local perturbations) and atmospheric effects cannot be separated.
+\item Observations of snowflake orientations show symmetrical distributions around 0$^\circ$ and standard deviations that confirm some of the assumptions made in theoretical studies. 
+\item Sheltered and unsheltered sites show systematic differences in orientation distributions, influenced by wind. Instrument (local perturbations) and atmospheric effects cannot be separated.
 \item Numerical simulations reveal that common estimation choices made in past studies may lead to data processing artifacts.
@@
 \begin{abstract}
-This study investigates the orientation of falling snowflakes using data from a Multi-Angle Snowflake Camera (MASC).
+Knowledge about orientation of falling snowflakes is still poorly documented with field measurements despite its importance, for example, in the interpretation of remote sensing data.
+This study investigates the orientation of  snowflakes using data from a Multi-Angle Snowflake Camera (MASC).
 We explore the impact of different observational setups (sheltered vs.
 unsheltered), wind conditions, hydrometeor type, and axis ratio on the orientation distributions.
-Numerical simulations are used to select the best orientation estimator and to understand the reason behind different results reported in past literature.
-We find that previously reported non-zero median orientations are likely artifacts of averaging absolute values of apparent orientations from individual cameras.
-Observed orientations generally follow a symmetrical distribution around 0$^\circ$, with broader and flatter distributions observed in unsheltered sites and/or high wind conditions.
+Numerical simulations are used to select the best orientation estimator and to understand the reason behind contrasting results reported in past literature.
+We find that previously reported non-zero median orientations are likely artifacts due to averaging absolute values of orientations from the three individual cameras.
+Observed orientations generally follow a symmetrical distribution around 0$^\circ$, with broader distributions observed at unsheltered sites and/or high wind conditions.
 Observed distributions may vary significantly from those assumed in previous studies, highlighting the need for further research on snowflake orientation under varying environmental conditions.
@@
 \section*{Plain Language Summary}
-Understanding the orientation of falling snowflakes is important for interpreting weather radar data and better understand snowfall micro-physics.
+Understanding the orientation of falling snowflakes is important for interpreting remote sensing data and better understand snowfall micro-physics.
 This study uses a 3-view camera to capture detailed images of falling snowflakes and estimate their orientations.
-Snowflakes generally fall with their largest dimension parallel to the ground (horizontal orientation), with variations in both directions around this setup.
-We found that observations in sites exposed to the wind show broader distributions of orientations compared to those in sheltered areas.
+Snowflakes generally fall with their largest dimension parallel to the ground, with typical variability of the orientation angle of about 30 degrees.
+We found that observations at sites exposed to the wind show broader distributions of orientations compared to those in sheltered areas.
 Additionally, our simulations suggest that some previous studies may have incorrectly reported average orientations because of the way data was processed.
-These findings help refine our knowledge of snowflake behavior and show the need of better understanding the physical mechanisms affecting the distribution of snowflake orientations.
+These findings help refine our understanding of snowflake behavior and show the need to better comprehend the physical mechanisms affecting the distribution of snowflake orientations.
 
@@
 
-Ice-phase hydrometeors, such as individual crystals and aggregates, fall with a preferred orientation.
-In stable conditions, they align in a way that maximizes the area projected in the direction of flow \cite{Koebschall_EF_2023}.
-This fact is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where.
+Ice-phase hydrometeors, such as individual crystals and aggregates, are typically assumed to fall with a preferred orientation.
+It is generally considered that crystals populating frozen clouds aligns preferentially in a way that maximizes the area projected in the vertical direction, as deduced by remote sensing and laboratory experiments.
+This was considered to be true for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}.
+This assumption is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where.
 Except for relatively rare situations, like storm electrification \cite{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology relies on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally.
-The concept of orientation is thus linked to the anisotropic nature of most hydrometeors and loses any meaning for isotropic, spherical particles.
+The concept of orientation is thus linked to the anisotropic geometry of most hydrometeors and loses any meaning for isotropic, spherical particles.
+In the case of snowflakes, they are often approximated by ellipsoids (or ellipses, in case of 2D images as shown in Fig.~\ref{fig:triplet}) and the orientation is thus defined by the angle between the horizontal and the major axis of the ellipsoid.
 
 Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ (i.e., horizontally oriented) values.
-Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly employ a Gaussian distribution model, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors.
+Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly assume a Gaussian distribution, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors.
 Observational studies \cite{Garrett_GRL_2015} confirm that differences exist due to hydrometeor type and environmental conditions (wind and turbulence).
-Observational studies report as well non-zero mean and median orientation values, which is contradictory with what was stated previously.
-As we will discuss later in this manuscript, this discrepancy is, in reality, the result of the way multi-view instrumental observations are combined to retrieve an individual estimate of orientation.
+However, they report as well non-zero mean and median orientation values, which is contradictory with what was stated previously.
+As we will discuss later in this study, this discrepancy is, in reality, the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation.
 
-The problem of observing hydrometeor orientations while they fall, without generating disturbance induced by the measurement itself, is not trivial.
+The problem of observing the orientations of hydrometeors while they fall, without generating disturbance induced by the measurement itself, is not trivial.
 Among the ground-based instruments designed to perform this task, we may cite the two-dimensional video disdrometer (2DVD) \cite{Kruger_JAOT_2002}, the multi-angle snowflake camera (MASC) \cite{Garrett_AMT_2012}, the Precipitation Imaging Package (PIP) \cite{Pettersen_atmosphere_2020}, and the Video In Situ Snowfall Sensor (VISSS) \cite{Maahn_AMT_2024}.
@@
 
-In this manuscript, we present an analysis of the fall orientation of ice-phase hydrometeors, conducted on a large dataset collected by a MASC.
-We explore the effect of different siting of the instrument (sheltered or unsheltered), wind conditions, hydrometeor type, and axis ratio, as well as the estimator of orientation used.
-For the last investigation, we complement the analysis with simulations aimed at uncovering biases that may emerge with some estimation choices.
+In this study, we present an analysis of the fall orientation of ice-phase hydrometeors, conducted on a large dataset collected by a MASC.
+We explore the effect of sheltered or unsheltered siting of the instrument, wind conditions, hydrometeor type, and axis ratio, as well as the estimator of orientation used.
+For the last investigation, we complement the analysis with simulations aimed at uncovering biases that may arise with some estimation choices.
 
@@
  \centering
- \begin{tabular}{l p{80mm}} 
+ \begin{tabular}{l p{90mm}} 
  \hline
@@
    \hline
-    \cite{Ryzhkov_JAMC_2011} & $\overline{\theta} = 0^\circ, \sigma = 40^\circ$ (aggregates) \newline 
-   $\overline{\theta} = 0^\circ \sigma = 40^\circ$ (graupel) \\
+    \cite{Ryzhkov_JAMC_2011} & $\overline{\theta} = 0^\circ, \sigma = 40^\circ$ (aggregates or graupel) \\
    \hline
@@
    \hline   
-   \cite{Putnam_MWR_2017} & $\overline{\theta} = 0^\circ \sigma = 20^\circ$ (snow) \newline 
-   $\overline{\theta} = 0^\circ \sigma \in [0,60]^\circ$ (graupel) \\
+   \cite{Putnam_MWR_2017} & $\overline{\theta} = 0^\circ , \sigma = 20^\circ$ (snow) \newline 
+   $\overline{\theta} = 0^\circ , \sigma \in [0,60]^\circ$ (graupel) \\
    \hline
-   \cite{Bukovic_JAMC_2018} &  $\overline{\theta} = 0^\circ, \sigma = 10^\circ$ (dendrites) \newline  $\overline{\theta} = 0^\circ \sigma = 40^\circ$ (aggregates) \\
+   \cite{Bukovic_JAMC_2018} &  $\overline{\theta} = 0^\circ, \sigma = 10^\circ$ (dendrites) \newline  $\overline{\theta} = 0^\circ , \sigma = 40^\circ$ (aggregates) \\
    \hline
-    \cite{Matsui_JGRA_2019}  & $\overline{\theta} = 0^\circ \sigma = 20^\circ$ (aggregates) \newline 
-   $\overline{\theta} = 20^\circ \sigma = 42^\circ$ (graupel) \\
+    \cite{Matsui_JGRA_2019}  & $\overline{\theta} = 0^\circ , \sigma = 20^\circ$ (aggregates) \newline 
+   $\overline{\theta} = 20^\circ , \sigma = 42^\circ$ (graupel) \\
    \hline
@@
    \cite{Gergely_JGRA_2016}$^*$ &  $\theta_{Q50} \in [38,47.3]^\circ$, $MAD(\theta) \in [43,51]^\circ$ \\
+   \hline
+   \cite{Stout_ACP_2024}$^*$ &  $\overline{\theta} \in [0,28]^\circ$ (various planar crystals, varying Re number)\\
    \hline
@@
    \hline
-    \multicolumn{2}{l}{$^{*}$Papers that present orientation statistics from ground-based measurements.}
+    \multicolumn{2}{l}{$^{*}$Papers that present orientation statistics based on field observations.}
  \end{tabular}
@@
 We utilize a large dataset of MASC observations, which includes co-located environmental information such as temperature, humidity, and wind, sourced from the MASCdb database extensively described in \cite{Grazioli_SD_2022}.
-The MASC instrument~\cite{Garrett_AMT_2012} captures simultaneous images of falling snowflakes with three cameras looking at a common focal point from three different viewpoints (0$^\circ$, +36$^\circ$, -36$^\circ$).
+The database includes observations collected in various sites worldwide, for example in the Swiss Alps, in Korea and in Antarctica between 2014 and 2024.
+The database includes both observations collected in conditions of natural wind exposure and, for some campaigns, with the instrument installed within a Double-Fence Intercomparison Reference~\cite<DFIR, see for example>{Smith_HESS_2020}, providing enhanced wind sheltering and protection from local turbulence and mean wind .
 
-MASCdb includes approximately one million image triplets (see Fig.~\ref{fig:triplet}) and pre-computed geometrical and microphysical descriptors.
-For the purpose of the present study, we filter the initial dataset by: (i) excluding precipitation occurring at positive temperatures, focusing thus on dry snow; (ii) removing particles classified as \textit{small particles} following the criteria in \cite{Praz_AMT_2017}, due to their size being too close to the image resolution (which is about 35$\mu$m); and (iii) excluding measurements potentially affected or dominated by wind-blown snow according to \cite{Schaer_TC_2020}.
+The MASC instrument captures simultaneous images of falling snowflakes with three coplanar cameras looking at a common focal point from three different viewpoints, with viewing angles of 0$^\circ$, +36$^\circ$, and -36$^\circ$.
+
+MASCdb includes approximately one million image triplets (see Fig.~\ref{fig:triplet} as example) and pre-computed geometrical and microphysical descriptors.
+For the purpose of the present study, we filter the initial dataset by: (i) excluding precipitation occurring at positive temperatures, as we focus on dry snow and not on melting hydrometeors or rain; (ii) removing particles classified as \textit{small particles} following the criteria in \cite{Praz_AMT_2017}, because of their size being too close to the image resolution (which is about 35$\mu$m) hampering the definition or estimation of their shape and orientation; and (iii) excluding measurements potentially affected or dominated by wind-blown snow according to \cite{Schaer_TC_2020}, as the focus of the study is on precipitation (snowfall) and not horizontal transport.
 All filtering steps are conducted on data already integrated into MASCdb, ensuring reproducible results.
@@
 
-In handling three 2D images of an individual 3D object, a primary challenge is to derive a single  estimate of orientation for each snowflake.
+When handling three 2D images of an individual 3D object, a primary challenge is to derive a single  estimate of orientation for each snowflake.
 Snowflakes are commonly modeled as ellipsoids or spheroids in previous MASC studies: the orientation of the major axis of the best-fitting ellipsoid in each 2D image is estimated, and the average of the three views is considered as a proxy for orientation \cite{Garrett_GRL_2015, Gergely_JGRA_2016, Fitch_AMT_2021, Fitch_JGR_2022}.
@@
 Examining Table~\ref{table:orientations}, the experimental studies using these estimators report the occurrence of non-zero modal orientation values between 10$^\circ$ and 20$^\circ$, challenging the common assumption of a distribution centered at 0$^\circ$.
-While the non-zero modal values reported for \textit{Ellipsoid\_Fit} in \cite{Jiang_JAS_2019} are based on a small sample of aggregate snowflakes, making evaluation difficult, we demonstrate that the appearance of non-zero modal values in the \textit{Mean\_Abs} approach results simply from averaging absolute values.
-This will offer a reconciliation between the common assumptions made on snowflake orientations and the contradiction of the first studies that used the MASC, a contradiction that was observed also by~\citeA{Garrett_GRL_2015} and discussed in the Conclusions of their works.
+While the non-zero modal values reported for \textit{Ellipsoid\_Fit} in \cite{Jiang_JAS_2019} are based on a small sample of aggregate snowflakes, making evaluation difficult, we demonstrate that the appearance of non-zero modal values in the \textit{Mean\_Abs} approach simply results from averaging absolute values.
+This will offer a reconciliation between the common assumptions made on snowflake orientations and the contradiction of the first studies that used the MASC, a contradiction that was also noted by~\citeA{Garrett_GRL_2015} and discussed in the Conclusions of their work.
 
 To address this, and to select an appropriate orientation estimator, we conducted numerical simulations (results shown in Fig.~\ref{fig:simulations}).
-These simulations generated ellipsoids with randomly varying geometries, as well as more realistic and geometrically complex aggregates following methods such as those described in \cite{Leinonen_AMT_2021}.
+These simulations generated ellipsoids with randomly varying geometries, as well as more realistic and geometrically complex aggregates following the same method as \citeA{Leinonen_AMT_2021}.
 The objects, of known 3D structure are projected into MASC-like 2D images to test various estimation methods.
 The orientation of the reference snowflake or ellipsoids is known and drawn from a Gaussian distribution with a zero mean and 25$^\circ$ standard deviation.
-Although this simulation setup could be expanded extensively to investigate the contribution of each ingredient at play, the goal here is to support the selection of an estimator to use with actual field-collected data.
+Although this simulation setup could be expanded extensively to investigate the contribution of each ingredient at play, the goal here is only to support the selection of an estimator to use with actual field-collected data.
 Figure~\ref{fig:simulations} provides an overview of how various methods reproduce the orientation distribution (panels a and c) and the magnitude of errors relative to the absolute orientation value (panels b and d).
@@
 (a) and (c): distribution of orientation value, both reference and reconstructed.
-(b) and (d): letter-value plots of the distribution of the error of the estimation of the absolute value of the orientation.
+(b) and (d): letter-value plots for the  error distribution of the orientation (in absolute value, not sign dependent).
 The simulations are conducted on 10'000 ellipsoids of various geometry as well as on 10'000 aggregates.
@@
 
-The result of the simulations reveal that the non-zero modal values observed with \textit{Mean\_Abs} are artifact resulting from averaging absolute values,  reducing the likelihood of near-zero estimates.
+The result of the simulations reveal that the non-zero modal values observed with \textit{Mean\_Abs} are artifacts resulting from averaging absolute values, reducing the likelihood of near-zero estimates.
 With an example: let us assume that the individual estimate of orientation from the three cams were -10$^\circ$, 0$^\circ$ and +20$^\circ$.
-With \textit{Mean\_Abs} the final estimate would be approximately 10$^\circ$, while for \textit{Mean} it would be roughly 3$^\circ$.
-The results shown here show that, at least statistically-speaking, a bias is introduced.
+With \textit{Mean\_Abs} the final estimate would be 10$^\circ$, while for \textit{Mean} it would be 3.33$^\circ$.
+The results presented here show that, at least statistically-speaking, a bias is introduced.
 
@@
 Individual camera views yield larger errors and broader distributions, which is in itself a noteworthy result given that many snowflake imagers only capture one view.
-\textit{Min} provides low errors but produces distributions that are excessively leptokurtic.
-Consequently, for subsequent analyses, we adopt a straightforward \textit{Mean}, which is possibly an optimal balance between preserving distribution shape with minimal errors and simplicity.
+\textit{Min} provides low errors but produces distributions that are excessively leptokurtic (standardized kurtosis in excess of 5).
+Consequently, for subsequent analyses, we adopt a straightforward \textit{Mean}, which is possibly an optimal balance between preserving distribution shape, minimal errors and simplicity.
 Importantly, \textit{Mean} is the only method in this specific experimental setup that likely maintains Gaussianity,  according to a Kolmogorov-Smirnov test.
@@
 However, the distribution has fat tails with the frequency of large orientation values exceeding the ones of a Gaussian of equivalent standard deviation.
-The installation conditions of the instruments significantly influence this distribution: sites within Double-Fence Intercomparison Reference~\cite<DFIR, see for example>{Smith_HESS_2020}, providing enhanced wind sheltering and protection from local turbulence, show less pronounced tails.
-In contrast, unsheltered sites exhibit more pronounced tails and a flatter distribution near 0$^\circ$, ranging between -25$^\circ$ and 25$^\circ$.
+The installation conditions of the instruments significantly influence this distribution.
+DFIR sites show less pronounced tails; in contrast, sites without this artificial sheltering exhibit more pronounced tails and a flatter distribution near 0$^\circ$, ranging between -25$^\circ$ and 25$^\circ$.
 This suggests that unsheltered observations are likely perturbed by local turbulence induced by the instrument itself, as described in~\citeA{Fitch_AMT_2021}.
@@
 One hypothesis could be that until 4 ms$^{-1}$ the DFIR is effective in minimizing local turbulence (DFIR curve constant, unsheltered curve abruptly increasing), while it gradually loses this capacity until "saturating" around 10 ms$^{-1}$ (DFIR curve increasing, unsheltered curve constant).
-Making this assumption is however implying that all the observed variations is due to the instrument itself and not a physical phenomenon of interaction of turbulent air and snowflakes.
+Making this assumption is however implying that all the observed variations are due to the instrument itself and not a physical phenomenon of interaction of turbulent air and snowflakes.
 Further research is therefore necessary to clarify these dynamics.
@@
 Given that the measurements within DFIRs are less perturbed, we focus now on this siting type.
-With the original dataset's large size, we conduct further data stratification.
+With the original dataset's large size, we can conduct further data stratification and still keep a large data sample.
 Figure~\ref{fig:hydroclass} illustrates orientation distributions under two contrasting wind regimes ($<4$ ms$^{-1}$ and $>8$ ms$^{-1}$), stratified by hydrometeor type~\cite{Praz_AMT_2017} and average axis ratio.
-Stronger winds lead to broader and platykurtic distributions, where a clear median around 0$^\circ$ orientation values is less evident and the distribution starts to assume uniform shape.
+Stronger winds lead to broader and flatter distributions, where a clear mode around 0$^\circ$ orientation values is less evident.
 For elongated columnar and planar crystals (low axis ratio), this results in distributions showing off-zero modal values, which were not observed under low-wind conditions.
+A possible explanation is the increasing mean shear with higher wind velocity, which may induce the rotation of suspended particles~\cite{Saffman_JFM_1956}.
 
-The influence of axis ratio on distribution is not significant, but this is an important information.
+The influence of axis ratio on distribution is not significant, but this is also an important information.
 It allows to exclude that the observed distribution could be an artefact given by the convention of assigning 0$^\circ$ orientation values to particles with axis ratio of 1.
-As the shape of the distribution is preserved for low and high axis ratios, we can be confident of the physical nature of this distribution shape.
+As the shape of the distribution is preserved for low and high axis ratios, we can be confident in the physical nature of this distribution shape.
  Focusing on low-wind conditions ($<4$ ms$^{-1}$) within DFIR sites, i.e.
@@
 \section{Conclusions}
-The main results of this study on the estimation of the orientation of falling snowflakes using a MASC are summarized as follows.
-At first, estimating orientation as the mean of the orientation of the best-fit ellipsoid from each camera view represents a reasonable approach in terms of minimizing errors and accurately reproducing the distribution of orientation values.
+This study focused on the estimation of the orientation of falling snowflakes using a MASC, providing observational back-up to commonly made assumptions on orientation distributions.
+It showed how some common data processing choices (for the data processing of MASC data) can lead to biased estimates, and highlighted the effect of unsheltered sites on the collected measurements.
+The main results are summarized as follows. 
+At first, estimating orientation as the mean of the orientation of the best-fit ellipsoid from each camera view represents a reasonable approach in terms of minimizing errors and accurately reproducing the distribution of (3D) orientation values.
 Single-camera estimates exhibit high uncertainty compared to combining estimates from all three views and this should be kept in mind if single-view imagers are used.
-Previous studies reporting median orientations deviating from 0$^\circ$ in MASC studies are likely artifacts resulting from averaging the \textit{absolute value} of orientation estimates from each camera view.
-This explication, demonstrated with numerical simulations in the manuscript, is simpler than what proposed for example by~\cite{Garrett_GRL_2015} who hypothesized fluttering and local, instrument-induced, turbulence (which has anyway an impact) as possible explanations.
+Previous studies reporting median orientations deviating from 0$^\circ$ in MASC studies are likely erroneous because of the artificially non-zero mode resulting from averaging the \textit{absolute value} of orientation estimates from each camera view.
+This explanation, supported by numerical simulations in the present study, is simpler than what proposed for example by~\cite{Garrett_GRL_2015} who hypothesized fluttering and local, instrument-induced, turbulence (which has anyway an impact) as possible explanations.
 Field measurements from a large dataset demonstrate that orientation values generally exhibit a symmetrical distribution around 0$^\circ$.
 Observations from unsheltered sites or in very high wind conditions result in broader and flatter orientation distributions, with a less pronounced mode at 0$^\circ$.
-This is likely influenced by both local disturbances (instrument-induced) and environmental factors, although with the current information we can only observe this discrepancy.
+This is likely influenced by both local disturbances (instrument-induced) and environmental factors, although with the current information we cannot further decipher the two sources.
 It is clear that further research, with different setups is needed to separate the contributions of these effects and estimate the \textit{natural} effect of wind and turbulence on the distribution of orientation values.
+For this purpose, instruments that do not create local perturbations (like PIP) could be used, ideally co-located with a MASC.
 Finally, we observed that orientation distributions of some hydrometeor types can vary significantly with respect to the assumptions made in previous studies.
@@
 The MASCDB dataset used to for the field-based data of this study is openly available in~\citeA{MASCDB_dataset_2023}.
-The scripts to reproduce the results in this manuscript are openly available at \hl{TODO: add github repo with the scripts}.
+The scripts to reproduce the results in this manuscript are openly available at this public library:~\url{https://zenodo.org/records/13941866} as well as with codes integrated in the pymascdb package:~\url{https://doi.org/10.5281/zenodo.7398285}.
 
@@
 \acknowledgments
+Jacopo Grazioli has received financial support from the Swiss National Science Foundation (grant no. 200020\_175700/1).
 

### 2024-12-01 Review_R1_answers_202412.tex
Updates from Overleaf

@@
 The distribution of wind components likely plays a more significant role.}
+
+    We agree with the reviewer that the distribution of wind components and turbulence at the relevant scales is probably more important than simple wind speed that we could investigate in this work (as all the MASC measurements in the database have wind speed information).
+
+    We carefully revised the manuscript to make sure not to overstate the role of wind speed alone and also to clarify when we cannot separate between the "natural" contribution of wind on the distribution of orientations and the "artificial" contribution due to the turbulence generated by strong wind interacting with the hollow measurement system of the MASC.
+A few changes are highlighted in the manuscript, we can cite for example in Sec.~3:
+
+    \blue{Atmospheric conditions are expected to play a role in shaping orientation distributions: atmospheric turbulence is probably a key driver, as well as horizontal winds interacting with the instrument frame~\cite{Fitch_AMT_2021}.
+}
+
+    and later in the same section:
+
+    \blue{This interpretation cannot unfortunately tell us any relevant information about the role of environmental turbulence, but only the likely effect of local turbulence introduced by the instrument.
+Further research is therefore necessary to clarify these dynamics, in particular the role of natural turbulence in broadening the orientation distribution.}
+
+    \hl{To beef up either the answer or the modification on the manuscript}
+
 \end{enumerate}
@@
     \item \textit{Including results on graupel in a study focused specifically on snowflakes raises questions about the sampling method.}
+
+    Our intention was not to focus only on snowflakes (meaning aggregate snowflakes) but on all the hydrometeors that can be reasonably observed with a MASC instrument.
+We realized that we were not particularly careful in the naming, and we revised the manuscript with care preferring expressions as "snow", "snow particles", "snow hydrometeors", "ice-phase hydrometeors" instead of "snowflakes".
+The title itself is rephrased as: \blue{Observation of the orientation  of snow hydrometeors at sheltered and unsheltered sites}.
+
+    We are however aware that retrieving orientation from MASC images of very flat hydrometeors (plates, dendrites) is likely less accurate than retrieving it from aggregate snowflakes or graupel.
+This happens because planar crystals become recognizable and classifiable with the method of~\cite{Praz_AMT_2017}, used in this paper, only when their basal facet face reasonably the cameras.
+We commented about this in the manuscript as well:
+
+    \blue{The results, especially for planar crystal types, must be considered with extreme care.
+In fact, the classification method of~\cite{Praz_AMT_2017}, which is trained on a supervised set, is able to identify planar crystals when their basal facet is facing towards the MASC cameras, making them recognizable.
+This introduces a bias as some orientations would not allow for classification.
+For example, a thin plate falling perfectly horizontal, would appear in MASC images as a thin column.
+}
 \end{enumerate}
@@
 However, I see the topic itself as highly relevant for active and passive microwave remote sensing but also for estimation of ice particle sedimentation velocity.
-Therefore, I recommend the manuscript for publication in GRL.}
+Therefore, I recommend the manuscript for publication in GRL.}\\
+
+\noindent
+We thank the reviewer for their positive feedback and for the many insightful suggestions.
+We tried to implement most of their comments in the revised version of the manuscript, and we addressed the specific comments here below.
 
@@
 
-    We now modified the manuscript to give some space to commenting and elaborating on the findings of~\cite{Stout_ACP_2024}. At first, we include this study in Table~1.
+    We now modified the manuscript to give some space to commenting and elaborating on the findings of~\cite{Stout_ACP_2024}.
+At first, we include this study in Table~1.
+Then, in the Introduction, we mention this study as:\\
+    \blue{ A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$.
+}\\
+
+    And in the Conclusions:\\
+    \blue{It must be noted that, at least in quiescent conditions and for some types of planar crystals, spiralling falling modes around a non-zero angle have been observed in the laboratory~\cite{Stout_ACP_2024} but it remains to be seen if this behavior can be observed and isolated also in natural conditions.
+With this respect, our database is not refined enough to be able to isolate efficiently individual types of planar crystals.
+}\\
 
@@
 
-    \item Introduction: Honestly, comparing it with the other sections, I find the introduction a bit weak.
+    \item \textit{Introduction: Honestly, comparing it with the other sections, I find the introduction a bit weak.
 Considering the wider audience addressed by GRL, I think the importance of ice particle orientation should be emphasized and explained more.
@@
 So characterizing the particle orientation is quite key to simulate its terminal velocity (which is quite an important parameter in any cloud model).
-Ice particle orientation is also relevant for determining cloud radiative properties such as optical depth or albedo.
+Ice particle orientation is also relevant for determining cloud radiative properties such as optical depth or albedo.}
+
+    We thank the reviewer for the suggestion.
+While we left the second part of the Introduction a bit more "specialized" (targeting an audience interested in snow imagers and their data processing), we realized that we did not provide a satasfactory broad overview, neglecting all the implications listed by the reviewer.
+We rephrased therefore the first part of the Introduction, removed a few technical paragraphs and included most of the suggestions of the reviewer.
+Some new paragraphs now read:
+
+    \blue{The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where.
+This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite{Eyre_QJRMS_2022}.
+For example, recent studies showed that considering hydrometeor orientation in the assimilation of high frequency radiometry may improve the  forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities as the Ice Water Path~\cite{Kaur_RS_2022}.}
+
+    \blue{The orientation and falling regime also affects the settling velocity of ice phase hydrometers, a key parameter linking microphysical properties to other tangible quantities as the precipitation intensity, and that remains to be better understood and documented~\cite{Heymsfield_JAS_2004}.}\\
+
+    \hl{I think the reviewer is right and we could make the Introduction even more broad and accessible. I did the changes above but suggestions welcome}
 
@@
 Do we have solid knowledge about their falling orientation?
-Also the references you provide a few sentences later (\cite{Noel_JAMC_2005,Tinklenberg_JFM_2023}) focus only on planar crystals.}\\
+Also the references you provide a few sentences later (\cite{Noel_JAMC_2005,Tinklenberg_JFM_2023}) focus only on planar crystals.}
+
+    We believe the reviewer is right.
+We were biased tp think mostly in terms of larger snow particles (snow aggregates, graupel) or planar crystals, so our formulation was overstating the current knowledge.
+We rephrased fully this sentence as:
+    \blue{Ice-phase hydrometeors are typically assumed to fall with a preferred orientation.
+It is generally considered that falling snow particles populating frozen clouds align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
+Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes.
+}\\
 
@@
 
-    In this case our intention was to differentiate between individual crystals (the shape of which can indeed be complicated, but that usually have a clear symmetry) from aggregates and graupel, which are composed by many crystals with additional contribution of riming, making it hard to understand the individual shapes or define a clear geometry.
-Our phrasing was confusing and this sentence needed to be reformulated in any case (see previous points).
-It now reads:
+    We agree that this sentence needed to be rephrased. It now reads:
 
-     \hl{TO DO}
+    \blue{ It is generally considered that falling snow particles populating frozen clouds align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
+Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}.
+}\\
 
@@
 
-    \cite{Klett_JAS_1995, Reinking_JAMC_1997}
+    This is a good point.
+We did not find studies providing explicitly quantitative values of orientation distributions for columns and needles (which would be a very interesting topic to tackle).
+There are a few important work that deal with the orientation of columns/needles, such as~\cite{Klett_JAS_1995, Reinking_JAMC_1997, Melnikov_JAOT_2013}, They however investigate the theoretical effect of turbulence on orientation (Klett) or the effect of varying orientation on remote sensing observations (Reinking, Melnikov) rather than exploring or making hypothesis about the "nature" of orientation of columns and needles.
 
@@
 
-    \item L.
+    \item \textit{L.
 211-215: One general thought regarding the fact that radar studies most often "retrieve" smaller range of orientation angles: If one looks at vertically pointing radar images, one always finds the lowest few hundred meters much more turbulent than the ice and snow cloud aloft.
@@
 So I am wondering if the often observable low turbulence in many ice clouds (for example indicated by small spectrum width) could be the reason why the orientation angle distribution is more narrow?
-Of course, the discrepancy might also be connected to unrealistic scattering properties often assumed by polarimetric shape retrievals.
+Of course, the discrepancy might also be connected to unrealistic scattering properties often assumed by polarimetric shape retrievals.}
 
-    \item Connected to this topic but even more complicated to answer is why some polarimetric radar observations seem to indicate a certain non-zero distribution of the particle orientation: \cite{Melnikov_JAOT_2013}
+    This would be indeed a good research question.
+In order to answer to it, ideally it would be needed to collect MASC data (or data with a similar imager) also in the snow clouds aloft the surface.
+There are significant advances in this technology (imagers carried by tethered balloons for example) but we are not there yet.
+Airborne imagers (on research aircrafts), for which many data exist, are not able to preserve the natural orientation of the particles in their measurements.
+To add to the complexity, radar data themselves are often perturbed by ground clutter, they need a few hundred meters to form their beam or have internal hardware limitations (like TR-limiters) that affect short range measurements.
+Intuitively we can expect that, with respect to "quieter" high level clouds, near the surface we expect a broader variability driven by turbulence generated by surface friction.
+
+    \hl{See if we can add something}
+
+    \item \textit{Connected to this topic but even more complicated to answer is why some polarimetric radar observations seem to indicate a certain non-zero distribution of the particle orientation: \cite{Melnikov_JAOT_2013}}
+
+    The work of~\cite{Melnikov_JAOT_2013} presents some retrievals of "fluttering intensity" from polarimetric radar data, being this intensity the standard deviation of the  distribution of orientation angles of ice-phase hydrometeors (which they assume  0-centered in their work, with zero being horizontal, although using the interval [$0,\pi$] in their definition).
+A possible source of confusion is that they mention sentences as "the median $\sigma_{\Theta}$ lies in an interval from 9 to 15\dots", but it is indeed the median of the standard deviations and not the median of the orientations.
+This study thus agrees with the other ones presented in our Table~1 and with the "standard" assumption of preferential horizontal orientation, at least in statistical sense.
+We decided to included this work in Table~1 and we thank the reviewer for the suggestion.
+
 \end{enumerate}

### 2024-12-01 Grazioli_JGRA_2024_R1.tex
Updates from Overleaf

@@
 
-\title{Observation of falling snowflakes orientation at sheltered and unsheltered sites}
+\title{Observation of the orientation  of snow hydrometeors at sheltered and unsheltered sites}
 
@@
 \item Observations of snowflake orientations show symmetrical distributions around 0$^\circ$ and standard deviations that confirm some of the assumptions made in theoretical studies. 
-\item Sheltered and unsheltered sites show systematic differences in orientation distributions, influenced by wind. Instrument (local perturbations) and atmospheric effects cannot be separated.
+\item Sheltered and unsheltered sites show systematic differences in orientation distributions. Instrument (local perturbations) and atmospheric effects cannot be separated.
 \item Numerical simulations reveal that common estimation choices made in past studies may lead to data processing artifacts.
@@
 \begin{abstract}
-Knowledge about orientation of falling snowflakes is still poorly documented with field measurements despite its importance, for example, in the interpretation of remote sensing data.
-This study investigates the orientation of  snowflakes using data from a Multi-Angle Snowflake Camera (MASC).
+Knowledge about orientation of falling snow is still poorly documented with field measurements despite its importance, for example, in the interpretation of remote sensing data.
+This study investigates the orientation of snow hydrometeors using data from a Multi-Angle Snowflake Camera (MASC).
 We explore the impact of different observational setups (sheltered vs.
-unsheltered), wind conditions, hydrometeor type, and axis ratio on the orientation distributions.
+unsheltered), wind speed, hydrometeor type, and axis ratio on the orientation distributions.
 Numerical simulations are used to select the best orientation estimator and to understand the reason behind contrasting results reported in past literature.
@@
 Observed orientations generally follow a symmetrical distribution around 0$^\circ$, with broader distributions observed at unsheltered sites and/or high wind conditions.
-Observed distributions may vary significantly from those assumed in previous studies, highlighting the need for further research on snowflake orientation under varying environmental conditions.
+Observed distributions may vary significantly from those assumed in previous studies, highlighting the need for further research on hydrometeor orientations under varying environmental conditions.
 \end{abstract}
@@
 \section*{Plain Language Summary}
-Understanding the orientation of falling snowflakes is important for interpreting remote sensing data and better understand snowfall micro-physics.
-This study uses a 3-view camera to capture detailed images of falling snowflakes and estimate their orientations.
-Snowflakes generally fall with their largest dimension parallel to the ground, with typical variability of the orientation angle of about 30 degrees.
+Understanding the orientation of falling snow particles is important for interpreting remote sensing data and better understand snowfall micro-physics.
+This study uses a 3-view camera to capture detailed images of falling particles and estimate their orientations.
+They generally fall with their largest dimension parallel to the ground, with typical variability of the orientation angle of about 30 degrees.
 We found that observations at sites exposed to the wind show broader distributions of orientations compared to those in sheltered areas.
 Additionally, our simulations suggest that some previous studies may have incorrectly reported average orientations because of the way data was processed.
-These findings help refine our understanding of snowflake behavior and show the need to better comprehend the physical mechanisms affecting the distribution of snowflake orientations.
+These findings help refine our understanding of snowfall behavior and show the need to better comprehend the physical mechanisms affecting the distribution of orientations.
 
 \section{Introduction}
+Ice-phase hydrometeors are typically assumed to fall with a preferred orientation.
+It is generally considered that falling snow particles populating frozen clouds align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
+Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes.
 
-Ice-phase hydrometeors, such as individual crystals and aggregates, are typically assumed to fall with a preferred orientation.
-It is generally considered that crystals populating frozen clouds aligns preferentially in a way that maximizes the area projected in the vertical direction, as deduced by remote sensing and laboratory experiments.
-This was considered to be true for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}.
-This assumption is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where.
-Except for relatively rare situations, like storm electrification \cite{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology relies on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally.
-The concept of orientation is thus linked to the anisotropic geometry of most hydrometeors and loses any meaning for isotropic, spherical particles.
-In the case of snowflakes, they are often approximated by ellipsoids (or ellipses, in case of 2D images as shown in Fig.~\ref{fig:triplet}) and the orientation is thus defined by the angle between the horizontal and the major axis of the ellipsoid.
+The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where.
+This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite<e.g.,>{Eyre_QJRMS_2022}.
+For example, recent studies showed that considering hydrometeor orientation in the assimilation of high frequency radiometry may improve the  forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities as the Ice Water Path~\cite{Kaur_RS_2022}.
+
+The orientation and falling regime also affects the settling velocity of ice phase hydrometers, a key parameter linking microphysical properties to other tangible quantities as the precipitation intensity, and that remains to be better understood and documented~\cite{Heymsfield_JAS_2004}.
+%Except for relatively rare situations, like storm electrification \cite{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology relies on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally.
+The concept of orientation is linked to the anisotropic geometry of most snow particles and loses any meaning for isotropic, spherical particles.
+In the case of snowfall, the hydrometeors are often approximated by ellipsoids (or ellipses, in case of 2D images as shown in Fig.~\ref{fig:triplet}) and the orientation is thus defined by the angle between the horizontal and the major axis of the ellipsoid.
 
@@
 Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly assume a Gaussian distribution, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors.
-Observational studies \cite{Garrett_GRL_2015} confirm that differences exist due to hydrometeor type and environmental conditions (wind and turbulence).
-However, they report as well non-zero mean and median orientation values, which is contradictory with what was stated previously.
-As we will discuss later in this study, this discrepancy is, in reality, the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation.
+Observational studies conducted in the field \cite{Garrett_GRL_2015} confirm that differences exist due to hydrometeor type and environmental conditions (wind and turbulence).
+They report as well non-zero mean and median orientation values, which is contradictory with what was stated previously.
+As we will discuss later in this study, this discrepancy can be the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation, at least for more complex three-dimensional snow particles.
+A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$.
+A study with similar methods, but focussing on larger snow aggregates showed instead that horizontal orientation is preferred for those hydrometeors~\cite{Koebschall_EF_2023}.
 
-The problem of observing the orientations of hydrometeors while they fall, without generating disturbance induced by the measurement itself, is not trivial.
+The problem of observing the orientations of hydrometeors while they fall in natural outdoor conditions, without generating disturbance induced by the measurement itself, is not trivial.
 Among the ground-based instruments designed to perform this task, we may cite the two-dimensional video disdrometer (2DVD) \cite{Kruger_JAOT_2002}, the multi-angle snowflake camera (MASC) \cite{Garrett_AMT_2012}, the Precipitation Imaging Package (PIP) \cite{Pettersen_atmosphere_2020}, and the Video In Situ Snowfall Sensor (VISSS) \cite{Maahn_AMT_2024}.
@@
    \hline
-    \cite{Kennedy_JAMC_2011}  &   $\overline{\theta} = 0^\circ, \sigma = 15^\circ$ (dendrites) \newline $\overline{\theta} = 0^\circ, \sigma = 30^\circ$ (aggregates)     \\
+    \cite{Kennedy_JAMC_2011}  &   $\overline{\theta} = 0^\circ, \sigma = 15^\circ$ (dendrites), 30$^\circ$ (aggregates) \\
+   \hline
+    \cite{Melnikov_JAOT_2013} & $\overline{\theta} = 0^\circ, \sigma  \in [2,20]^\circ$ (ice-phase clouds) \\
    \hline
@@
    \hline
-   \cite{Bukovic_JAMC_2018} &  $\overline{\theta} = 0^\circ, \sigma = 10^\circ$ (dendrites) \newline  $\overline{\theta} = 0^\circ , \sigma = 40^\circ$ (aggregates) \\
+   \cite{Bukovic_JAMC_2018} &  $\overline{\theta} = 0^\circ, \sigma = 10^\circ$ (dendrites) , 40$^\circ$ (aggregates) \\
    \hline
@@
    \hline
-   \cite{Stout_ACP_2024}$^*$ &  $\overline{\theta} \in [0,28]^\circ$ (various planar crystals, varying Re number)\\
+   \cite{Stout_ACP_2024}$^*$ &  $\overline{\theta} \in [0,28]^\circ$ (planar crystals, varying Re number)\\
    \hline
@@
    \hline
-   \cite{Fitch_JGR_2022}$^*$ & $\theta_{M} = 3^\circ$ 
-   \newline
-   $\theta_{M} = 13^\circ$ (different estimations)  \\
+   \cite{Fitch_JGR_2022}$^*$ & $\theta_{M} = 3^\circ$ or  13$^\circ$ (different estimations)  \\
    \hline
@@
 While the non-zero modal values reported for \textit{Ellipsoid\_Fit} in \cite{Jiang_JAS_2019} are based on a small sample of aggregate snowflakes, making evaluation difficult, we demonstrate that the appearance of non-zero modal values in the \textit{Mean\_Abs} approach simply results from averaging absolute values.
-This will offer a reconciliation between the common assumptions made on snowflake orientations and the contradiction of the first studies that used the MASC, a contradiction that was also noted by~\citeA{Garrett_GRL_2015} and discussed in the Conclusions of their work.
+This will offer a reconciliation between the common assumptions made on hydrometeor orientations and the contradiction of the first studies that used the MASC, a contradiction that was also noted by~\citeA{Garrett_GRL_2015} and discussed in the Conclusions of their work.
 
@@
 Consequently, for subsequent analyses, we adopt a straightforward \textit{Mean}, which is possibly an optimal balance between preserving distribution shape, minimal errors and simplicity.
-Importantly, \textit{Mean} is the only method in this specific experimental setup that likely maintains Gaussianity,  according to a Kolmogorov-Smirnov test.
+\textit{Mean} is the only method in this specific experimental setup that likely maintains Gaussianity,  according to a Kolmogorov-Smirnov test.
 
@@
 
-Atmospheric conditions are expected to play a role in shaping orientation distributions, particularly increasing wind conditions.
+Atmospheric conditions are expected to play a role in shaping orientation distributions: atmospheric turbulence is probably a key driver, as well as horizontal winds interacting with the instrument frame~\cite{Fitch_AMT_2021}.
 Figure~\ref{fig:wind-effect} compares the standard deviation of orientation distributions between DFIR and unsheltered sites at different wind speeds.
 At low wind speeds ($<1$ ms$^{-1}$) and high wind speeds ($>10$ ms$^{-1}$), the curves are close.
-However, they diverge by up to 5/6$^\circ$ for intermediate wind speeds, with the DFIR curve showing an increase with wind speed once the wind speed exceeds 4 ms$^{-1}$ and being relatively constant below that threshold.
-Under near-zero wind conditions, local turbulence induced by the instrument geometry is minimal, making DFIR and unsheltered conditions comparable.
+They diverge by up to 5/6$^\circ$ for intermediate wind speeds, with the DFIR curve showing an increase with wind speed once the wind speed exceeds 4 ms$^{-1}$ and being relatively constant below that threshold.
+Under near-zero wind conditions, local turbulence induced by the instrument  should be minimal, making DFIR and unsheltered conditions comparable.
 One hypothesis could be that until 4 ms$^{-1}$ the DFIR is effective in minimizing local turbulence (DFIR curve constant, unsheltered curve abruptly increasing), while it gradually loses this capacity until "saturating" around 10 ms$^{-1}$ (DFIR curve increasing, unsheltered curve constant).
-Making this assumption is however implying that all the observed variations are due to the instrument itself and not a physical phenomenon of interaction of turbulent air and snowflakes.
-Further research is therefore necessary to clarify these dynamics.
+This interpretation cannot unfortunately tell us any relevant information about the role of environmental turbulence, but only the likely effect of local turbulence introduced by the instrument.
+Further research is therefore necessary to clarify these dynamics, in particular the role of natural turbulence in broadening the orientation distribution.
 
@@
 A possible explanation is the increasing mean shear with higher wind velocity, which may induce the rotation of suspended particles~\cite{Saffman_JFM_1956}.
+The results, especially for planar crystal types, must be considered with extreme care.
+In fact, the classification method of~\cite{Praz_AMT_2017}, which is trained on a supervised set, is able to identify planar crystals when their basal facet is facing towards the MASC cameras, making them recognizable.
+This introduces a bias as some orientations would not allow for classification.
+For example, a thin plate falling perfectly horizontal, would appear in MASC images as a thin column.
 
@@
 Single-camera estimates exhibit high uncertainty compared to combining estimates from all three views and this should be kept in mind if single-view imagers are used.
-Previous studies reporting median orientations deviating from 0$^\circ$ in MASC studies are likely erroneous because of the artificially non-zero mode resulting from averaging the \textit{absolute value} of orientation estimates from each camera view.
+Previous studies reporting median orientations deviating from 0$^\circ$ in MASC studies (in natural outdoor conditions) are likely erroneous because of the artificially non-zero mode resulting from averaging the \textit{absolute value} of orientation estimates from each camera view.
 This explanation, supported by numerical simulations in the present study, is simpler than what proposed for example by~\cite{Garrett_GRL_2015} who hypothesized fluttering and local, instrument-induced, turbulence (which has anyway an impact) as possible explanations.
-Field measurements from a large dataset demonstrate that orientation values generally exhibit a symmetrical distribution around 0$^\circ$.
+It must be noted that, at least in quiescent conditions and for some types of planar crystals, spiralling falling modes around a non-zero angle have been observed in the laboratory~\cite{Stout_ACP_2024} but it remains to be seen if this behavior can be observed and isolated also in natural conditions.
+With this respect, our database is not refined enough to be able to isolate efficiently individual types of planar crystals.
+
+Field measurements from our large dataset demonstrate that orientation values generally exhibit a symmetrical distribution around 0$^\circ$.
 Observations from unsheltered sites or in very high wind conditions result in broader and flatter orientation distributions, with a less pronounced mode at 0$^\circ$.

### 2024-12-11 Review_R1_answers_202412.tex
Updates from Overleaf

@@
     \blue{Atmospheric conditions are expected to play a role in shaping orientation distributions: atmospheric turbulence is probably a key driver, as well as horizontal winds interacting with the instrument frame~\cite{Fitch_AMT_2021}.
-}
+Turbulence has been shown to heavily impact the falling of particles in fluids in general~\cite{Petersen_JFM_2019,Brandt_ARFM_2022} and of frozen hydrometeors in particular~\cite{Nemes_JFM_2017,Li_JFM_2021}.}
 
@@
 Further research is therefore necessary to clarify these dynamics, in particular the role of natural turbulence in broadening the orientation distribution.}
-
-    \hl{To beef up either the answer or the modification on the manuscript}
-
 \end{enumerate}
@@
 This is exactly what we showed, in our paper, to be due to the bias of averaging \textbf{absolute value} of orientation estimates taken from different viewpoints and it was one of the driving motivation of our publication.
+        \item A further limitation is that the study considers objects falling in a liquid, which implies a far lower density ratio compared to hydrometeors in air, which in turn can impact the settling dynamics as discussed in recent studies by one of our coauthors (see~\cite{Tinklenberg_JFM_2023,Tinklenberg_JFM_2024}).
     \end{itemize}
@@
     \blue{ A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$.
+A study with similar methods, but focussing on larger snow aggregates showed instead that horizontal orientation is preferred for those hydrometeors~\cite{Koebschall_EF_2023}.
+It is to be remarked that those and other studies, in considering objects falling in liquids, use a solid-to-fluid density ratio which is orders lower than in atmospheric precipitation.
+This may have far-reaching consequences in the settling dynamics~\cite{Tinklenberg_JFM_2023,Tinklenberg_JFM_2024}.
 }\\
@@
     \blue{The orientation and falling regime also affects the settling velocity of ice phase hydrometers, a key parameter linking microphysical properties to other tangible quantities as the precipitation intensity, and that remains to be better understood and documented~\cite{Heymsfield_JAS_2004}.}\\
-
-    \hl{I think the reviewer is right and we could make the Introduction even more broad and accessible. I did the changes above but suggestions welcome}
 

### 2024-12-11 Grazioli_JGRA_2024_R1.tex
Updates from Overleaf

@@
 A study with similar methods, but focussing on larger snow aggregates showed instead that horizontal orientation is preferred for those hydrometeors~\cite{Koebschall_EF_2023}.
+It is to be remarked that those and other studies, in considering objects falling in liquids, use a solid-to-fluid density ratio which is orders lower than in atmospheric precipitation.
+This may have far-reaching consequences in the settling dynamics~\cite{Tinklenberg_JFM_2023,Tinklenberg_JFM_2024}.
 
@@
 Atmospheric conditions are expected to play a role in shaping orientation distributions: atmospheric turbulence is probably a key driver, as well as horizontal winds interacting with the instrument frame~\cite{Fitch_AMT_2021}.
+Turbulence has been shown to heavily impact the falling of particles in fluids in general~\cite{Petersen_JFM_2019,Brandt_ARFM_2022} and of frozen hydrometeors in particular~\cite{Nemes_JFM_2017,Li_JFM_2021}.
+
  Figure~\ref{fig:wind-effect} compares the standard deviation of orientation distributions between DFIR and unsheltered sites at different wind speeds.

### 2024-12-11 Review_R1_answers_202412.tex
Updates from Overleaf

@@
 I am therefore returning the paper to you so that you can make the necessary changes.
-}
+}\\
+
+\noindent
+In this document we address point-by-point the remarks of the two reviewers and make the necessary changes in the manuscript.
 
@@
 The findings from the numerical experiment, together with the analysis of the in-situ data, will be a valuable addition to the literature and a useful resource for the remote sensing community.
-I recommend the manuscript be published in GRL, after the following issues are addressed:}
+I recommend the manuscript be published in GRL, after the following issues are addressed:}\\
+
+\noindent
+We thank the Reviewer for their positive feedback and for the suggestions and comments listed below.
 
@@
 
-    We agree with the reviewer that the distribution of wind components and turbulence at the relevant scales is probably more important than simple wind speed that we could investigate in this work (as all the MASC measurements in the database have wind speed information).
+    We agree with the reviewer that the distribution of wind components and turbulence at the relevant scales is probably more important than simple wind speed that we could investigate in this work (as all the MASC measurements in the database have co-located wind speed information available).
 
     We carefully revised the manuscript to make sure not to overstate the role of wind speed alone and also to clarify when we cannot separate between the "natural" contribution of wind on the distribution of orientations and the "artificial" contribution due to the turbulence generated by strong wind interacting with the hollow measurement system of the MASC.
-A few changes are highlighted in the manuscript, we can cite for example in Sec.~3:
+A few changes are highlighted in the manuscript.
+We may cite for example in Sec.~3:
 
@@
 
-    We thank the reviewer for this, extremely relevant, recommendation.
+    We thank the reviewer for this relevant, recommendation.
 This high quality study, with laboratory experiments using snowflake replicas in a liquid solution shows that planar crystals may be falling with regimes that do not have a preferential orientation around 0$^\circ$.
@@
     \begin{itemize}
-        \item A relevant limitation of this work is that "it  only considers quiescent conditions, as the authors want to know under what conditions particle instability still occurs, even without the addition of turbulence.".
+        \item A relevant limitation of~\cite{Stout_ACP_2024} work is that, quoting, "it  only considers quiescent conditions, as the authors want to know under what conditions particle instability still occurs, even without the addition of turbulence.".
 This is quite different with respect to our field-based experiments.
@@
 Then, in the Introduction, we mention this study as:\\
+
     \blue{ A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$.
-A study with similar methods, but focussing on larger snow aggregates showed instead that horizontal orientation is preferred for those hydrometeors~\cite{Koebschall_EF_2023}.
+A study with similar methods, but focused on larger snow aggregates, instead showed that horizontal orientation is preferred for these hydrometeors~\cite{Koebschall_EF_2023}.
 It is to be remarked that those and other studies, in considering objects falling in liquids, use a solid-to-fluid density ratio which is orders lower than in atmospheric precipitation.
@@
 This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite{Eyre_QJRMS_2022}.
-For example, recent studies showed that considering hydrometeor orientation in the assimilation of high frequency radiometry may improve the  forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities as the Ice Water Path~\cite{Kaur_RS_2022}.}
+For example, recent studies showed that considering hydrometeor orientation in the assimilation of high-frequency radiometry may improve forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities such as the Ice Water Path~\cite{Kaur_RS_2022}.
+}
 
@@
 We were biased tp think mostly in terms of larger snow particles (snow aggregates, graupel) or planar crystals, so our formulation was overstating the current knowledge.
-We rephrased fully this sentence as:
+We rephrased fully this sentence as:\\
+
     \blue{Ice-phase hydrometeors are typically assumed to fall with a preferred orientation.
-It is generally considered that falling snow particles populating frozen clouds align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
+Falling snow particles populating frozen clouds are generally considered to align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
 Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes.
@@
 
-    We agree that this sentence needed to be rephrased. It now reads:
-
-    \blue{ It is generally considered that falling snow particles populating frozen clouds align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
-Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}.
-}\\
+    We agree that this sentence needed to be rephrased (see previous point).
 
@@
 
-    \hl{See if we can add something}
-
     \item \textit{Connected to this topic but even more complicated to answer is why some polarimetric radar observations seem to indicate a certain non-zero distribution of the particle orientation: \cite{Melnikov_JAOT_2013}}

### 2024-12-11 Grazioli_JGRA_2024_R1.tex
Updates from Overleaf

@@
 Ice-phase hydrometeors are typically assumed to fall with a preferred orientation.
-It is generally considered that falling snow particles populating frozen clouds align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
+Falling snow particles populating frozen clouds are generally considered to align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
 Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes.
@@
 This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite<e.g.,>{Eyre_QJRMS_2022}.
-For example, recent studies showed that considering hydrometeor orientation in the assimilation of high frequency radiometry may improve the  forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities as the Ice Water Path~\cite{Kaur_RS_2022}.
+For example, recent studies showed that considering hydrometeor orientation in the assimilation of high-frequency radiometry may improve forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities such as the Ice Water Path~\cite{Kaur_RS_2022}.
 
-The orientation and falling regime also affects the settling velocity of ice phase hydrometers, a key parameter linking microphysical properties to other tangible quantities as the precipitation intensity, and that remains to be better understood and documented~\cite{Heymsfield_JAS_2004}.
+The orientation and falling regime also affects the settling velocity of ice phase hydrometeors, a key parameter linking microphysical properties to other tangible quantities as the precipitation intensity, and that remains to be better understood and documented~\cite{Heymsfield_JAS_2004}.
 %Except for relatively rare situations, like storm electrification \cite{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology relies on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally.
@@
 
-Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ (i.e., horizontally oriented) values.
+Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ values (i.e., horizontally oriented).
 Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly assume a Gaussian distribution, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors.
-Observational studies conducted in the field \cite{Garrett_GRL_2015} confirm that differences exist due to hydrometeor type and environmental conditions (wind and turbulence).
-They report as well non-zero mean and median orientation values, which is contradictory with what was stated previously.
+Observational studies conducted in the field \cite{Garrett_GRL_2015} confirm that there are differences due to the type of hydrometeor and environmental conditions (wind and turbulence).
+They also report non-zero mean and median orientation values, which is contradictory with what was stated previously.
 As we will discuss later in this study, this discrepancy can be the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation, at least for more complex three-dimensional snow particles.
 A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$.
-A study with similar methods, but focussing on larger snow aggregates showed instead that horizontal orientation is preferred for those hydrometeors~\cite{Koebschall_EF_2023}.
+A study with similar methods, but focused on larger snow aggregates, instead showed that horizontal orientation is preferred for these hydrometeors~\cite{Koebschall_EF_2023}.
 It is to be remarked that those and other studies, in considering objects falling in liquids, use a solid-to-fluid density ratio which is orders lower than in atmospheric precipitation.
@@
 These sensors allow for direct microphysical investigations based on pictures of individual hydrometeors.
-As mentioned above, indirect retrievals of the behavior of large populations of particles within the sampling volume of remote-sensing instruments are also commonly employed, mostly by means of dual-polarization technologies \cite{Matrosov_JAS_2005}.
+As mentioned above, indirect retrievals of the behavior of large populations of particles within the sampling volume of remote sensing instruments are also commonly used, mostly by means of dual-polarization technologies \cite{Matrosov_JAS_2005}.
 
@@
 
-We utilize a large dataset of MASC observations, which includes co-located environmental information such as temperature, humidity, and wind, sourced from the MASCdb database extensively described in \cite{Grazioli_SD_2022}.
+We used a large data set of MASC observations, which includes co-located environmental information such as temperature, humidity, and wind, sourced from the MASCdb database extensively described in \cite{Grazioli_SD_2022}.
 The database includes observations collected in various sites worldwide, for example in the Swiss Alps, in Korea and in Antarctica between 2014 and 2024.
@@
 
-MASCdb includes approximately one million image triplets (see Fig.~\ref{fig:triplet} as example) and pre-computed geometrical and microphysical descriptors.
-For the purpose of the present study, we filter the initial dataset by: (i) excluding precipitation occurring at positive temperatures, as we focus on dry snow and not on melting hydrometeors or rain; (ii) removing particles classified as \textit{small particles} following the criteria in \cite{Praz_AMT_2017}, because of their size being too close to the image resolution (which is about 35$\mu$m) hampering the definition or estimation of their shape and orientation; and (iii) excluding measurements potentially affected or dominated by wind-blown snow according to \cite{Schaer_TC_2020}, as the focus of the study is on precipitation (snowfall) and not horizontal transport.
+MASCdb includes approximately one million image triplets (see Fig.~\ref{fig:triplet} as an example) and precomputed geometrical and microphysical descriptors.
+For the purpose of the study, we filter the initial dataset by: (i) excluding precipitation occurring at positive temperatures, as we focus on dry snow and not on melting hydrometeors or rain; (ii) removing particles classified as \textit{small particles} following the criteria in \cite{Praz_AMT_2017}, because of their size being too close to the image resolution (which is about 35$\mu$m) hampering the definition or estimation of their shape and orientation; and (iii) excluding measurements potentially affected or dominated by wind-blown snow according to \cite{Schaer_TC_2020}, as the focus of the study is on precipitation (snowfall) and not horizontal transport.
 All filtering steps are conducted on data already integrated into MASCdb, ensuring reproducible results.
@@
 For elongated columnar and planar crystals (low axis ratio), this results in distributions showing off-zero modal values, which were not observed under low-wind conditions.
-A possible explanation is the increasing mean shear with higher wind velocity, which may induce the rotation of suspended particles~\cite{Saffman_JFM_1956}.
+A possible explanation is the increasing mean shear with higher wind velocity, which may induce the rotation of suspended particles.
 The results, especially for planar crystal types, must be considered with extreme care.
@@
 This explanation, supported by numerical simulations in the present study, is simpler than what proposed for example by~\cite{Garrett_GRL_2015} who hypothesized fluttering and local, instrument-induced, turbulence (which has anyway an impact) as possible explanations.
-It must be noted that, at least in quiescent conditions and for some types of planar crystals, spiralling falling modes around a non-zero angle have been observed in the laboratory~\cite{Stout_ACP_2024} but it remains to be seen if this behavior can be observed and isolated also in natural conditions.
+It must be noted that, at least in quiescent conditions and for some types of planar crystals, spiraling falling modes around a non-zero angle have been observed in the laboratory~\cite{Stout_ACP_2024} but it remains to be seen if this behavior can be observed and isolated also in natural conditions.
 With this respect, our database is not refined enough to be able to isolate efficiently individual types of planar crystals.

### 2024-12-16 Review_R1_answers_202412.tex
Updates from Overleaf

@@
 
-    and later in the same section:
+     Later in the same section:
 
@@
 Further research is therefore necessary to clarify these dynamics, in particular the role of natural turbulence in broadening the orientation distribution.}
+
+    And in the Conclusions:
+
+    \blue{It is clear that further research, with different setups is needed to separate the contributions of these effects and estimate the \textit{natural} effect of wind and turbulence on the distribution of orientation values.
+For this purpose, instruments that do not create local perturbations (like PIP) could be used, ideally co-located with a MASC.
+}
 \end{enumerate}
@@
     \blue{The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where.
-This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite{Eyre_QJRMS_2022}.
+This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite{Eyre_QJRMS_2022}.
 For example, recent studies showed that considering hydrometeor orientation in the assimilation of high-frequency radiometry may improve forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities such as the Ice Water Path~\cite{Kaur_RS_2022}.
@@
     \blue{Ice-phase hydrometeors are typically assumed to fall with a preferred orientation.
-Falling snow particles populating frozen clouds are generally considered to align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
+Falling snow particles populating frozen clouds are  considered to align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
 Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes.

### 2024-12-16 Grazioli_JGRA_2024_R1.tex
Updates from Overleaf

@@
 Ice-phase hydrometeors are typically assumed to fall with a preferred orientation.
-Falling snow particles populating frozen clouds are generally considered to align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
+Falling snow particles populating frozen clouds are considered to align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.
 Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes.
@@
 The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where.
-This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite<e.g.,>{Eyre_QJRMS_2022}.
+This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite<e.g.,>{Eyre_QJRMS_2022}.
 For example, recent studies showed that considering hydrometeor orientation in the assimilation of high-frequency radiometry may improve forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities such as the Ice Water Path~\cite{Kaur_RS_2022}.
@@
 
-Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ values (i.e., horizontally oriented).
+Considering a population of hydrometeors rather than individual particles, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ values (i.e., horizontally oriented).
 Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly assume a Gaussian distribution, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors.
@@
 As we will discuss later in this study, this discrepancy can be the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation, at least for more complex three-dimensional snow particles.
-A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$.
+A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiraling fall regime with preferential orientations that depart from 0$^\circ$.
 A study with similar methods, but focused on larger snow aggregates, instead showed that horizontal orientation is preferred for these hydrometeors~\cite{Koebschall_EF_2023}.
